---
title: "Engine Test: Julia"
format:
  html:
    code-fold: false
    code-tools:
        source: true
        toggle: true
        caption: none
engine: julia
julia:
    exeflags: ["--project=@."]
filters:
    - show-lang.lua # show language in code block
cache: true
---


Trying this on Julia engine to overcome limitations of knitr. Want to be able to pass objects between all three languages. The trouble so far has been that while it is easy to run R and Python chunks from the Julia engine, I have not been able to get either to run from the proper .venv or renv, or to recognize the packages in those environments.

## Setup

```{julia}
println("Working directory:", pwd())
```

```{julia}

# ENV["PYTHON"] = "../.venv/Scripts/python.exe"
# ENV["PYTHON"] = "/c/Users/cdonov12/local/repos/SMdocs/.venv/Scripts/python.exe"
# ENV["JULIA_PYTHONCALL_EXE"] = "../.venv/Scripts/python.exe"
# ENV["JULIA_PYTHONCALL_EXE"] = "../.venv/Scripts/python.exe"
# ENV["JULIA_PYTHONCALL_EXE"] = "/c/Users/cdonov12/local/repos/SMdocs/.venv/Scripts/python.exe"
# ENV["JULIA_CONDAPKG_BACKEND"] = "Null"

using Pkg
using RCall
using PythonCall
```

### Julia

Check working directory

```{julia}
println("Working directory:", pwd())
```

Check active project

```{julia}
Base.active_project()
```

### R

Check working directory

```{r}
print(getwd())
```

Check where R is running from

```{r}
Sys.which("R")
```

### Python

```{python}
import os
print(os.getcwd())
```

```{python}
import sys
print(sys.executable)
```

```{python}
# import seaborn as sns
```

Can't get CallPython to use same environment