---
title: "Mix Python and R"
format:
  html:
    code-fold: true
---

To mix both R and python in same document, do not specify engine in yaml header. This will use knitr by default, and run reticulate for the python cells.

Potential problem is that the version of R that reticulate uses may not be the same as the one in the `.venv`. README has instructions on how to make sure reticulate uses the `.venv`. 

Also note a problem trying to run a python cell in a mixed doc interactively. Something hinky with matplotlib here it doesn't have access to a proper GUI to show plots. Works fine if previewed or rendered though.

## R Cell

```{r}
#| label: r_test
plot(iris)
```

Check directory:

```{r}
getwd()
```

Looks like all `.qmd` files default to project directory. In contrast `.ipynb` files default to the directory where it is located.

## Python Cell

Test a citation [@adriaanse1993].

```{python}
#| label: fig-polar
#| fig-cap: "A line plot on a polar axis"

import numpy as np
import matplotlib.pyplot as plt

r = np.arange(0, 2, 0.01)
theta = 2 * np.pi * r
fig, ax = plt.subplots(
  subplot_kw = {'projection': 'polar'} 
)
ax.plot(theta, r)
ax.set_rticks([0.5, 1, 1.5, 2])
ax.grid(True)
plt.show()
```