---
title: "Sustainability Metrics<br>Secondary Data"
author: "Chris Donovan"
institute: "Food Systems Research Center"
date: '2025-03-19'
date-format: long
format:
  revealjs:
    embed-resources: false
    incremental: false
    theme: [serif, preso.scss]
    transition: fade
    scrollable: true
    chalkboard: true
    revealjs-plugins:
      - revealjs-text-resizer
editor: 
  markdown: 
    wrap: 72
bibliography: ../fsrc.bib
citations-hover: true
citeproc: true
include-in-header:
 - text: |
     <style>
     #title-slide .title {
       font-size: 2em;
     }
     </style>
---

```{r}
#| label: prep
#| include: false
pacman::p_load(
  ggplot2,
  ggpubr,
  plotly,
  forcats,
  reactable,
  htmltools,
  stringr,
  Hmisc,
  reshape,
  performance,
  AER,
  sandwich,
  sjPlot,
  caret,
  ranger,
  glmnet,
  conflicted,
  knitr,
  kableExtra,
  sf,
  leaflet,
  ggraph,
  igraph,
  readr,
  RColorBrewer,
  paletteer,
  fmsb,
  purrr
)

source('dev/get_reactable.R')
source('dev/get_vt_spiders.R')

conflicted::conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  tidyr::expand(),
  .quiet = TRUE
)
```

## Introduction {.smaller}

:::: {.columns .v-center-container}

::: {.column width = '40%'}

-   Goals for secondary data:
    -   Identify existing data and gaps
    -   Explore methods of aggregating data
-   Structure of presentation:
    1.  Framework and methods
    2.  Preliminary results of case study
    3.  Uncertainty and sensitivity
-   Guiding questions:
    -   How well do the data represesent the system?
    -   Valuation, scaling, weighting, aggregation
    -   Where and how do we incorporate qualitative data?
    
:::

::: {.column width = '60%'}

![Intervale Farm, Sally McCay, UVM
Photo](../images/intervale_small.png){fig-align="center" width="90%"
fig-alt="A picture of a field at Intervale Farm."}

:::

::::

::: notes

-   The point is to take stock of existing data, swap out and fill in with primary research where necessary. 
-   Skeleton of a framework.

:::

# 1. Framework and Methods {.centered-section}

## Case Study: Validation by Cronbach {.smaller}

Raw + min max scaling + geometric aggregation

<!-- :::: {.columns style='display: flex !important; height: 80%;'} -->

<!-- ::: {.column width = "40%" style = 'display: flex; align-items: center;'} -->

:::: {.columns .v-center-container}

::: {.column width = '50%'}

-   Cronbach's Alpha is a measure of internal reliability or consistency
    [@cronbach1951CoefficientAlphaInternal]
-   Arbitrary guideline of 0.7 to validate a measurement tool
    [@taber2018UseCronbachAlpha]
\begin{equation}
\alpha=\frac{n}{n-1}\left( 1-\frac{\sum_{i}^{V_i}}{V_t} \right)
\end{equation}
Where $n$ is the number of items, $V_t$ is the variance of item $i$, and
$V_t$ is the variance of all items
-   It is not supposed to be negative.

:::

::: {.column width = '50%'}

```{r}
#| class: centered-vertical
cronbachs <- readRDS('preso/data/cronbach_dimensions.rds')
get_reactable(
  cronbachs,
  fullWidth = FALSE,
  searchable = FALSE,
  filterable = FALSE
)
```

:::

::::

::: {.notes}

-   Context for Cronbach: not necessarily what we want, just one way to show consistency within dimension. 
-   Might be okay if dimension is scattered

:::

# References
