---
title: "Regional Analyses"
format:
  html:
    fig-dpi: 200
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

::: {.callout-note collapse='false' title='Caution'}
The Sustainability Metrics project, as well as this site itself, are works in progress. All data and analyses shown here are preliminary. 
:::

```{r}
#| label: setup
#| echo: false
#| cache: false
pacman::p_load(
  dplyr,
  conflicted,
  purrr,
  stringr,
  ggraph,
  igraph,
  dplyr,
  readr,
  RColorBrewer,
  paletteer,
  snakecase,
  tidyr,
  htmltools,
  reactable,
  downloadthis
)
pacman::p_load_gh('ChrisDonovan307/projecter')
source('dev/get_dimension_ggraph.R')
sm_load()
conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::arrange(),
  dplyr::summarize(),
  dplyr::as_data_frame(),
  .quiet = TRUE
)
```


## Introduction {#sec-intro}

This group of pages will present a survey of secondary data at the regional/county level in the Northeast. Add more context here eventually.

## Dimension Graphs

```{r}
#| label: prep_graphs
#| include: false
# Filter DF and check data
source('dev/get_reactable.R')
tree <- data_paper_tree %>% 
  mutate(
    indicator = case_when(
        str_detect(indicator, 'inputs bought locally') ~ '% local inputs',
        str_detect(indicator, 'production inputs') ~ 'production inputs',
        str_detect(indicator, 'metabolic') ~ 'presence of disease',
        .default = indicator
      ) %>% 
      str_remove(' of food business')
  )
get_reactable(tree)

# Set up list of results
plots <- list()
```

### Economics

```{r}
#| fig-height: 5
#| fig-width: 15
plots$econ <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'economics',
  y_limits = c(-3.75, 3.25),
  palette = "scico::batlowW"
)
plots$econ
```

### Environment

```{r}
#| fig-height: 7
#| fig-width: 10
plots$env <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'environment',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$env
```

### Health

```{r}
#| fig-height: 8
#| fig-width: 10
plots$health <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'health',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$health
```

### Production

```{r}
#| fig-height: 5
#| fig-width: 12
plots$prod <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'production',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$prod
```

### Social

```{r}
#| fig-height: 6
#| fig-width: 13
plots$social <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'social',
  y_limits = c(-3, 3.25),
  palette = "scico::batlowW"
)
plots$social
```

```{r}
#| label: save_pngs
#| include: false
# df of dimensions to use

iwalk(plots, ~ {
  ggsave(
    paste0('temp/c3/', .y, '.png'),
    plot = .x,
    width = 11,
    height = 8,
    units = 'in',
    bg = 'white'
  )  
})
```

## Dimension Table

```{r}
data_paper_tree %>% 
  mutate(metric = case_when(
    str_detect(metric, 'NONE') ~ NA,
    .default = metric
  )) %>% 
  arrange(dimension, index, indicator, metric) %>% 
  get_reactable()
```

