---
title: "County Trends"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  caret,
  knitr,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  sf,
  skimr,
  ggplot2,
  ggpubr
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')

# Load SMdata package from repos folder
sm_load()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```

```{r}
#| label: load
#| include: false
#| eval: false

# THIS IS OLD
# Load metrics
sm_data <- readRDS('data/sm_data.rds')
metrics <- sm_data$metrics
meta <- sm_data$metadata

# Load frame
frame <- readRDS('data/frameworks/new_frame.rds')
```

## Wrangling

Load up new round of metrics from `SMdata` using the `sm_load()` function.

```{r}
#| output: false
# Check out summary from OneDrive that has our variable names
get_str(data_paper_meta)

# Pull vars to reduce metrics down
meta <- data_paper_meta %>% 
  filter(
    !is.na(variable_name),
    !variable_name %in% c('tbd', '', 'NONE'),
    resolution == 'county'
  ) %>% 
  select(
    variable_name,
    starts_with('n_'),
    latest_year,
    dimension:metric,
    weighting:desirable,
  )
get_str(meta)

# Also using weighting variables from summary table, exported in SMdata
get_str(weighting)

# Load metrics, reduce to neast counties, grab only relevant vars and weights
get_str(metrics)
dat <- metrics %>% 
  filter_fips('new') %>% 
  filter(variable_name %in% c(meta$variable_name, weighting$variable_name)) %>% 
  mutate(across(c(year, value), as.numeric))
get_str(dat)

# Pivot wider for transformations, then alphabetize columns
wide <- dat %>% 
  pivot_wider(
    id_cols = c(fips, year),
    values_from = 'value',
    names_from = 'variable_name'
  ) %>% 
  select(fips, year, sort(names(.)[2:length(names(.))]))
get_str(wide)
```

## Missing Data

```{r}
#| label: missing_data
#| output: false
skimr::skim(wide)
# Missing years are county areas, that's fine
# Other missing values are because years don't match up, that's fine

## Check a couple vars to make sure they are somewhat intact
get_str(wide)

# check expenses per farm expPF
get_missing(wide, 'expPF')

# check income from agritourism incomeAgTourismRecPropTotal
out <- get_missing(wide, 'incomeAgTourismRecPropTotal', TRUE)
get_str(out)
```

Make a table of missing data for each var

```{r}
#| label: check_miss
vars <- names(wide)[!names(wide) %in% c(
  'landAcres', 
  'waterAcres', 
  'fips', 
  'year'
)]
miss_table <- map(vars, \(name) {
  out <- wide %>% 
    select(fips, year, !!sym(name)) %>% 
    na.omit() %>% 
    complete(fips, year)
  data.frame(
    var = name,
    n_miss = sum(is.na(out[[name]])),
    total = length(out[[name]])
  ) %>% 
    mutate(
      prop_miss = round(n_miss/total, 2)
    )
}) %>% 
  bind_rows() %>% 
  arrange(desc(prop_miss))
```

```{r}
#| label: miss_table
get_reactable(miss_table)
```

## Explore Trends

```{r}
#| label: prep_df
#| output: false
get_str(dat)

# Remove weighting variables
dat <- dat %>% 
  filter(!variable_name %in% weighting$variable_name)
get_str(dat)

# Get names of vars that have > 1 time point
series_vars <- dat %>% 
  group_by(variable_name) %>% 
  summarize(n_year = length(unique(year))) %>% 
  filter(n_year > 1) %>% 
  pull(variable_name)
(len <- length(series_vars))
# 44 have more than 1 time point

# Filter to just those variables so we can do regressions
series_metrics <- filter(dat, variable_name %in% series_vars)
get_str(series_metrics)
```

`{r} len` metrics have > 1 data point at county level. 

### Time Series

Plot time series of average variable that has at least 2 data points

```{r}
#| label: graphs
#| fig-height: 25
#| fig-width: 12
# get_str(series_metrics)
# get_str(series_vars)

plots <- map(series_vars, ~ {
  df <- series_metrics %>% 
    filter(variable_name == .x) %>% 
    ggplot(aes(x = year, y = value)) + 
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) + 
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      # y = .x,
      title = .x
    ) +
    theme_classic()
})
# plots
# plots[[1]]

ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```

```{r}
#| label: for_c3
#| include: false
#| eval: false
# Pulling out a couple of graphs for c3 preso
get_str(dat)
c3_vars <- c(
  'poorMentalHealthDays',
  'frequentPhysicalDistress',
  'expPF',
  'lifeExpectancy',
  'acresPF',
  'foodInsecurityRate'
)

# Make a plot for each one
c3_plots <- map(c3_vars, ~ {
  df <- series_metrics %>% 
    filter(variable_name == .x) %>% 
    ggplot(aes(x = year, y = value)) + 
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) + 
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      y = ''
      # title = .x
    ) +
    expand_limits(y = 0) +
    theme_classic()
}) %>% 
  setNames(c(c3_vars))
c3_plots
c3_plots[[1]]

iwalk(c3_plots, ~ {
  ggsave(
    plot = .x,
    filename = paste0('temp/c3/c3_', .y, '.png'),
    height = 4,
    width = 6,
    units = 'in',
    dpi = 300
  )
})
```

## Linear Models

First trying linear models with time as independent variable.

```{r}
#| label: trend_analysis
#| output: false
res <- map(series_vars, ~ {
  tryCatch(
    {
      df <- series_metrics %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year)
        )
      model <- lm(value ~ year, data = df)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(series_vars)) %>% 
  discard(\(x) is.null(x))
```

Now that we have regression outputs, let's put them in a table to see which ones were significant. 

```{r}
#| label: make_table
#| output: false
# Put together one table of results from all models, add stars
res_df <- imap(res, ~ {
  .x %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many metrics
n_metrics <- nrow(res_df)

# How many are significant
perc_sig <- mean(res_df$sig == '*') * 100
```

We have `{r} n_metrics` metrics, `{r} round(perc_sig, 1)` of which vary significantly over time.

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch.

```{r}
#| label: make_sense_of_trends
#| output: false

# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
model_outputs <- res_df %>% 
  # filter(sig == '*') %>%
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(data_paper_meta)
directions <- data_paper_meta %>% 
  filter(desirable %in% c('higher', 'lower'), !is.na(variable_name)) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction. Then, 
model_outputs <- model_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
percs <- prop.table(tab) * 100
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse.

Check out table with outputs for each regression. Filter by significance and outcome. Trend is just whether it is going up or down, outcome is whether that is good or bad.

```{r}
#| label: model_table
get_reactable(
  model_outputs, 
  defaultPageSize = 10,
  columns = list(
   term = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
  
```

## Spatial Panel Regressions

Run spatial regression for all vars. Might be worth looking into whether there is a version of Moran test that can be run across years. Until then, just running each one with spatial error model. 

### Setup

```{r}
#| label: splm
#| output: false
#| eval: false
get_str(series_metrics)

dat <- series_metrics %>% 
  filter(
    variable_name != 'population',
    str_length(fips) == 5
  ) %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )

# Get rid of geometry, make our df a plm object
df <- dat %>% 
  select(fips, year, rentMedianPercHH) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  na.omit() %>% 
  mutate(rentMedianPercHH = as.numeric(rentMedianPercHH)) %>% 
  plm::pdata.frame()
get_str(df)

# Get our listw based on counties
lw <- neast_counties_2024 %>% 
  filter(fips %in% df$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Try FE error model with rentMedianPercHH
fe_err <- spml(
  rentMedianPercHH ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'within'
)
summary(fe_err)

# RE error model
re_err <- spml(
  rentMedianPercHH ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'random'
)
summary(re_err)

# Check to use RE or FE, see whether x_it related to a_i (idiosyncratic error)
test <- sphtest(fe_err, re_err)
test
# Not sig: use RE, more efficient

# Check RE output
summary(re_err)
```

### Run Models

So the workflow for each variable is:

1. Make sure panel is balanced, i.e. see if Connecticut will be a problem. Adjust accordingly.
2. Create list weight object based on the counties that are represented by in the variable. I think we have to do this for each variable individually... should double check though.
3. Run FE and RE model with year as numeric.
4. Hausman test with FE and RE model. If sig, report FE, if not, report RE

Try it:

```{r}
#| label: spregs_over_vars
#| output: false
#| eval: false
get_str(dat)

# Make variables numeric
dat <- dat %>% 
  mutate(across(3:last_col(), as.numeric))
get_str(dat)

# Get vector of vars
vars <- names(dat)[!names(dat) %in% c('fips', 'year')]

# Map over each var and do steps 1 through 5
# plan(multisession, cores = availableCores(omit = 2))
out <- map(vars, \(var) {
  
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var)) %>% 
      drop_na()
    
    # 1. Only use fips that produce balanced panel
    balanced_fips <- df %>% 
      group_by(fips) %>% 
      filter(n() == length(unique(df$year))) %>% 
      pull(fips) %>% 
      unique()
    
    df <- df %>% 
      filter(fips %in% balanced_fips)
    
    # 2. Create lw object
    lw <- neast_counties_2024 %>% 
      filter(fips %in% balanced_fips) %>% 
      # filter(fips %in% df$fips) %>% 
      poly2nb() %>% 
      nb2listw(zero.policy = TRUE)
    cat('\nCreated lw')
    
    # Make df a plm object
    df <- plm::pdata.frame(df)
    
    # 3. Run FE and RE
    formula <- as.formula(paste(var, "~ as.numeric(year)"))
    fe <- spml(
      formula,
      data = df, 
      listw = lw,
      model = 'within'
    )
    re <- spml(
      formula,
      data = df, 
      listw = lw,
      model = 'random'
    )
    cat('\nRan models')
    
    # 4. Hausman test. If sig, use FE
    test <- sphtest(fe, re)
    if (test$p.value < 0.05) {
      model <- fe
    } else {
      model <- re
    }
    cat('\nRan Hausman')
    
    return(list('model' = model, 'fips' = balanced_fips))
      
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
  
}, .progress = FALSE) %>% 
 setNames(c(vars))
# plan(sequential)

# Save for posterity
saveRDS(out, 'data/objects/spatial_regression_outs.rds')
```

Check how they turned out:

```{r}
#| output: false
out <- readRDS('data/objects/spatial_regression_outs.rds')

# How many were successful
perc_success <- mean(map(out, class) == 'list') * 100

get_str(out[[1]])
get_str(out)
out

# Check out errors only
errors <- out %>% 
  keep(is.character)
errors
error_vars <- names(errors)
```

`{r} round(perc_success, 1)`% of regressions were successful. Lots of metrics messed up because of uneven panels. They include: `{r} paste0(error_vars, collapse = ', ')`.

### Results

Check out the successful models:

```{r}
#| label: spatial_panel_table
model_outs <- out %>% 
  keep(is.list)
# get_str(model_outs)
# get_str(model_outs[[1]])
# get_str(model_outs[[1]][[1]])

# Check model type
# model_outs[[1]][[1]]$type
# table(map_chr(model_outs, ~ .x[[1]]$type))
# They are all RE models - no FE


## Test out outcomes
# sum <- summary(model_outs[[1]][[1]])
# coefs <- sum$CoefTable[2, ]

# Make a table
spatial_table <- imap(model_outs, ~ {
  tryCatch({
    
    # Pull model output results and coefficients
    sum <- summary(.x[[1]])
    coefs <- sum$CoefTable[2, ]
    
    df <- data.frame(
      n = length(.x$fips),
      type = .x[[1]]$type,
      var = .y,
      phi = sum$errcomp[1],
      rho = sum$errcomp[2],
      est = coefs[1],
      se = coefs[2],
      t = coefs[3],
      p = coefs[4]
    ) 
    
    # Figure out good or bad outcome
    # If sig, then check reverse. else blank
    # If reverse, reverse it, else same
    # If up, good, else bad
    df <- df %>% 
      mutate(
        sig = ifelse(p < 0.05, TRUE, FALSE),
        posneg = case_when(
          est > 0 ~ 'pos',
          est < 0 ~ 'neg',
          .default = ''
        ),
        reverse = ifelse(.y %in% reverse, TRUE, FALSE),
        direction = case_when(
          sig == TRUE & reverse == TRUE & posneg == 'pos' ~ 'neg',
          sig == TRUE & reverse == TRUE & posneg == 'neg' ~ 'pos',
          sig == TRUE & reverse == FALSE ~ as.character(posneg),
          .default = ''
        ),
        outcome = case_when(
          sig == FALSE ~ '',
          sig == TRUE & direction == 'pos' ~ 'better',
          sig == TRUE & direction == 'neg' ~ 'worse',
          .default = '-'
        )
      ) %>% 
      select(-c(posneg, direction))
    return(df)
  },
    error = function(e) {
      return(paste('Error:', e))
    }
  )
}) %>% 
  bind_rows()
# get_str(spatial_table)

# Round off and make table
spatial_table %>% 
  mutate(
    across(c(phi:p), ~ format(round(.x, 3), nsmall = 3)),
    type = case_when(
      str_detect(type, 'random') ~ 'RE',
      .default = NA
    )
  ) %>% 
  get_reactable(
    defaultColDef = colDef(minWidth = 50),
    columns = list(
      n = colDef(minWidth = 25),
      var = colDef(minWidth = 200),
      est = colDef(
        minWidth = 100,
        name = 'β'
      ),
      outcome = colDef(minWidth = 75),
      rho = colDef(name = 'ρ'),
      phi = colDef(name = 'φ'),
      p = colDef(name = 'p-value'),
      'p-value' = colDef(minWidth = 100)
    )
  )
```

- n: Number of counties included in regression
- type: FE = fixed effects (every county has $\beta$), RE = random effects (counties have different $\beta$s)
- var: Metric
- $\phi$: Spatial autocorrelation in idiosyncratic attributes (counties)
- $\rho$: Spatial autocorrelation in residual errors (omitted variables)
- $\beta$: Regression coefficient of year on the variable
- se: standard error of $\beta$
- t: test statistic
- p-value: p-value
- sig: whether p is < 0.05
- reverse: whether the metric has been reversed so that lower numbers are better and higher numbers are worse (like food insecurity)
- outcome: if the regression is significant, is it getting better or worse? Accounts for the 'reverse' variable
