---
title: "Metric Trends"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

::: {.callout-note collapse='false' title='Caution'}
This is a mildly slapdash page working out spatial regressions for time series metrics from the refined framework variables. There are some cleaning issues to address with some variables, and the scripts are wildly disorganized. 
:::

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  caret,
  knitr,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  sf,
  skimr
)


# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

# Load SMdata package from repos folder
sm_load()
```

```{r}
#| label: load
# Load metrics
sm_data <- readRDS('data/sm_data.rds')
metrics <- sm_data$metrics
meta <- sm_data$metadata

# Load frame
frame <- readRDS('data/frameworks/new_frame.rds')
```

This page explores what data we have available at the county level specifically. It is branching off of the main "Analysis" pages because they are state-level analyses and I want to leave it there for posterity.

Note that there are some cleaning issues with this dataset, and the scripts are wildly disorganized. The important thing is that we figured out how to iterate through a few dozen variables and run spatial error models at the county level to see if they are getting better or worse over time.

## Testing

Load up new round of metrics from `SMdata` using the `sm_load()` function.

```{r}
#| output: false
# Check out summary from OneDrive that has our variable names
get_str(data_paper_meta)

# Pull vars to reduce metrics down
meta <- data_paper_meta %>% 
  filter(
    !is.na(variable_name),
    !variable_name %in% c('tbd', '', 'NONE'),
    resolution == 'county'
  ) %>% 
  select(
    variable_name,
    starts_with('n_'),
    latest_year,
    dimension:metric,
    weighting:desirable,
  )
get_str(meta)

# Also using weighting variables from summary table, exported in SMdata
get_str(weighting)

# Load metrics, reduce to neast counties, grab only relevant vars and weights
get_str(metrics)
dat <- metrics %>% 
  filter_fips('new') %>% 
  filter(variable_name %in% c(meta$variable_name, weighting$variable_name)) %>% 
  mutate(across(c(year, value), as.numeric))
get_str(dat)

# Pivot wider for transformations, then alphabetize columns
wide <- dat %>% 
  pivot_wider(
    id_cols = c(fips, year),
    values_from = 'value',
    names_from = 'variable_name'
  ) %>% 
  select(fips, year, sort(names(.)[2:length(names(.))]))
get_str(wide)
```

### Missing Data

```{r}
#| label: missing_data
#| output: false
skimr::skim(wide)
# Missing years are county areas, that's fine
# Other missing values are because years don't match up, that's fine

## Check a couple vars to make sure they are somewhat intact
get_str(wide)

# check expenses per farm expPF
get_missing(wide, 'expPF')

# check income from agritourism incomeAgTourismRecPropTotal
out <- get_missing(wide, 'incomeAgTourismRecPropTotal', TRUE)
get_str(out)
```

Make a table of missing data for each var

```{r}
#| label: check_miss
vars <- names(wide)[!names(wide) %in% c(
  'landAcres', 
  'waterAcres', 
  'fips', 
  'year'
)]
miss_table <- map(vars, \(name) {
  out <- wide %>% 
    select(fips, year, !!sym(name)) %>% 
    na.omit() %>% 
    complete(fips, year)
  data.frame(
    var = name,
    n_miss = sum(is.na(out[[name]])),
    total = length(out[[name]])
  ) %>% 
    mutate(
      prop_miss = round(n_miss/total, 2)
    )
}) %>% 
  bind_rows() %>% 
  arrange(desc(prop_miss))
```

```{r}
#| label: miss_table
get_reactable(miss_table)
```

