---
title: "Regional Analyses"
format:
  html:
    fig-dpi: 200
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

::: {.callout-note collapse='false' title='Caution'}
The Sustainability Metrics project, as well as this site itself, are works in progress. All data and analyses shown here are preliminary. 
:::

```{r}
#| label: setup
#| echo: false
#| cache: false
pacman::p_load(
  dplyr,
  conflicted,
  purrr,
  stringr,
  ggraph,
  igraph,
  dplyr,
  readr,
  RColorBrewer,
  paletteer,
  snakecase,
  tidyr,
  htmltools,
  reactable,
  downloadthis
)
pacman::p_load_gh('ChrisDonovan307/projecter')
source('dev/get_dimension_ggraph.R')
sm_load()
conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::arrange(),
  dplyr::summarize(),
  dplyr::as_data_frame(),
  .quiet = TRUE
)
```


## Introduction {#sec-intro}

This group of pages will present a survey of secondary data at the regional/county level in the Northeast. Add more context here eventually.

## Dimension Graphs

```{r}
#| label: prep_graphs
#| include: false
# Filter DF and check data
source('dev/get_reactable.R')
tree <- data_paper_tree %>% 
  mutate(
    indicator = case_when(
        str_detect(indicator, 'inputs bought locally') ~ '% local inputs',
        str_detect(indicator, 'production inputs') ~ 'production inputs',
        str_detect(indicator, 'metabolic') ~ 'presence of disease',
        .default = indicator
      ) %>% 
      str_remove(' of food business')
  )
get_reactable(tree)

# Set up list of results
plots <- list()
```

### Economics

```{r}
#| fig-height: 5
#| fig-width: 15
plots$econ <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'economics',
  y_limits = c(-3.75, 3.25),
  palette = "scico::batlowW"
)
plots$econ
```

### Environment

```{r}
#| fig-height: 7
#| fig-width: 10
plots$env <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'environment',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$env
```

### Health

```{r}
#| fig-height: 8
#| fig-width: 10
plots$health <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'health',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$health
```

### Production

```{r}
#| fig-height: 5
#| fig-width: 12
plots$prod <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'production',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$prod
```

### Social

```{r}
#| fig-height: 6
#| fig-width: 13
plots$social <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'social',
  y_limits = c(-3, 3.25),
  palette = "scico::batlowW"
)
plots$social
```

## TEMP

```{r}
#| label: save_pngs
#| include: false
# df of dimensions to use
dims <- data.frame(
  heights = c(5, 7, 8, 5, 6),
  widths = c(15, 10, 10, 12, 13)
)
pwalk(list(plots, names(plots), 1:nrow(dims)), \(plot, name, i) {
  ggsave(
    paste0('temp/c3/', name, '.png'),
    plot = plot,
    height = dims[i, 1],
    width = dims[i, 2],
    units = 'in',
    bg = 'white',
    dpi = 500
  )  
})
```

```{r}
#| label: correlations
#| include: false
#| eval: false
pacman::p_load(
  Hmisc,
  tidyr,
  reshape2
)

get_str(metrics)
df <- metrics %>% 
  filter(variable_name %in% data_paper_meta$variable_name) %>% 
  get_latest_year() %>% 
  filter_fips('new') %>% 
  mutate(variable_name = str_split_i(variable_name, '_', 1)) %>%
  pivot_wider(
    id_cols = fips,
    values_from = value,
    names_from = variable_name
  )
get_str(df)
skimr::skim(df)

# Arrange metrics in order of dimensions
get_str(data_paper_meta)
order <- data_paper_meta %>% 
  arrange(dimension, variable_name) %>% 
  filter(variable_name %in% names(df)) %>% 
  pull(variable_name)
# check
data_paper_meta %>% 
  arrange(dimension, variable_name) %>% 
  filter(variable_name %in% names(df)) %>% 
  select(variable_name, dimension) %>% 
  group_by(dimension) %>% 
  summarize(count = n())

df <- df %>% 
  select(fips, any_of(order))
get_str(df)

mat <- df %>% 
  na.omit() %>% 
  select(-fips) %>% 
  as.matrix()

# Get correlations
cor <- rcorr(mat, type = 'pearson')

# Melt correlation values and rename columns
cor_r <- reshape::melt(cor$r) %>% 
  setNames(c('var_1', 'var_2', 'value'))

# Save p values
cor_p <- melt(cor$P)
p.value <- cor_p$value

# Make heatmap with custom text aesthetic for tooltip
corplot <- cor_r %>% 
  ggplot(aes(var_1, var_2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(
    low = "#762a83", 
    mid = "white", 
    high = "#1b7837", 
    midpoint = 0
  ) +
  geom_hline(yintercept = 14.5) +
  geom_hline(yintercept = 31.5) +
  geom_hline(yintercept = 45.5) +
  geom_hline(yintercept = 49.5) +
  geom_vline(xintercept = 14.5) +
  geom_vline(xintercept = 31.5) +
  geom_vline(xintercept = 45.5) +
  geom_vline(xintercept = 49.5) +
  theme(
    axis.text.x = element_text(
      hjust = 1, 
      angle = 45
    )
    # axis.text.x = element_blank(),
    # axis.ticks.x = element_blank()
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = NULL
  )
corplot
ggsave(
  plot = corplot,
  filename = 'temp/c3/corplot.png',
  height = 10,
  width = 12,
  dpi = 300
)
```

## Dimension Table

```{r}
data_paper_tree %>% 
  mutate(metric = case_when(
    str_detect(metric, 'NONE') ~ NA,
    .default = metric
  )) %>% 
  arrange(dimension, index, indicator, metric) %>% 
  get_reactable()
```

