---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  caret,
  knitr,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  spatialreg,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr,
  flextable
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')

# Load SMdata package from repos folder
sm_load()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

results <- list()
```

## Linear Models

First trying linear models with time as independent variable. Normalizing dataset before we do anything else.

```{r}
#| label: normalize
#| output: false
datasets <- readRDS('data/data_paper/datasets.rds')
get_str(datasets$dp_metrics_county)

# Function to min max then scale to 0 to 100 for percent change
min_max <- function(x) {
  normed <- (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  normed * 100
}

vars <- unique(datasets$dp_metrics_county$variable_name)
datasets$dp_metrics_county <- map(vars, ~ {
  datasets$dp_metrics_county %>%
    filter(variable_name == .x) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
get_str(datasets$dp_metrics_county)  
```

### Regressions

```{r}
#| label: trend_analysis
#| output: false
# Get a crosswalk to add state as a fixed effect
state_link <- fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

res <- map(datasets$series_vars, ~ {
  tryCatch(
    {
      df <- datasets$dp_metrics_county %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year),
          state_fips = str_sub(fips, end = 2)
        ) %>% 
        left_join(state_link, by = join_by(state_fips == fips))
      model <- lm(normed_value ~ year + state_code, data = df)
      return(model)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(datasets$series_vars)) %>%
  discard(\(x) is.null(x))
```

Now that we have regression outputs, let's put them in a table to see which ones were significant. 

```{r}
#| label: make_table
#| output: false
# Put together one table of results from all models, add stars
res_df <- imap(res, ~ {
  df <- .x %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
  conf_ints <- confint(.x)['year', ] %>% 
    as.vector()
  df$ci_low <- conf_ints[1]
  df$ci_high <- conf_ints[2]
  df
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many metrics
n_metrics <- nrow(res_df)

# How many are significant
perc_sig <- mean(res_df$sig == '*') * 100
```

We have `{r} n_metrics` metrics, `{r} round(perc_sig, 1)` of which vary significantly over time.

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch.

```{r}
#| label: make_sense_of_trends
#| output: false
# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
raw_outputs <- res_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(data_paper_meta)
directions <- data_paper_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose 4 metrics
model_outputs <- raw_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# See what we lost
lost <- setdiff(raw_outputs$term, model_outputs$term) %>% 
  paste0(collapse = ', ')

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
(percs <- prop.table(tab) * 100)
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} lost`).

Check out table with outputs for each regression. Filter by significance and outcome. Trend is just whether it is going up or down, outcome is whether that is good or bad.


### Table

```{r}
#| label: model_table
# Add framework info
framed_outputs <- data_paper_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(model_outputs, by = join_by(variable_name == term))

# Save some outputs to OneDrive
write_csv(
  framed_outputs,
  'outputs/lm_reg_outputs.csv'
)
flex <- framed_outputs %>% 
  select(-variable_name) %>% 
  as_flextable(show_coltype = FALSE)
flextable::save_as_docx(
  flex,
  path = 'outputs/lm_reg_outputs.docx'
)

get_reactable(
  framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```


### Figure

Figure below shows 95% confidence intervals for the coefficients of each metric, including state as a covariate.

```{r}
#| label: regression_figure
# get_str(framed_outputs)

out <- framed_outputs %>% 
  mutate(
    across(c(estimate, starts_with('ci')), as.numeric),
    outcome = ifelse(is.na(outcome), 'no change', outcome) %>% 
      snakecase::to_title_case()
  ) %>% 
  ggplot(aes(x = estimate, y = reorder(metric, estimate))) +
  geom_point(aes(color = outcome)) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
  geom_vline(xintercept = 0) +
  scale_color_manual(values = c('darkgreen', 'grey', 'purple')) +
  theme_minimal() +
  labs(
    x = 'Average Percent Change per Year',
    y = 'Metric',
    color = 'Outcome'
  )
ggsave(
  'outputs/lm_conf_ints.png',
  plot = out,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)
out
```


```{r}
#| label: get_weights
#| eval: false
#| include: false

### Weights section
# Adding weight variables eventually
get_str(datasets$dp_weights)
datasets$dp_weights$variable_name %>% 
  unique()
get_str(data_paper_meta)
data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name)

# Define weights crosswalk from metrics to variable names here
# TODO: Work this into data export from SMdata package
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5year',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)
weights_crosswalk

# Take our county data and join weight variable names to it
weights_crosswalk <- data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric)
  # mutate(weight_variable = ifelse(is.na(weight_variable, 'none',)))
get_str(weights_crosswalk)
weights_crosswalk

# Get weight for each var, given fips and year. This is what we use to join to
# our county metrics
get_str(datasets$dp_weights)
weights_by_year_fips <- datasets$dp_weights %>% 
  rename(
    weight_variable = variable_name,
    weight_value = value
  ) %>% 
  left_join(weights_crosswalk) %>% 
  drop_na(year) %>% 
  mutate(across(c(year, weight_value), ~ as.numeric(.x)))
weights_by_year_fips
get_str(weights_by_year_fips)
head(weights_by_year_fips, 100)
tail(weights_by_year_fips, 100)
# Unweighted vars have NAs, that's good. We will just drop them in join

# Now join to county metrics
get_str(weights_by_year_fips)
get_str(datasets$dp_metrics_county)
dat <- datasets$dp_metrics_county %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(fips, year, variable_name) %>% 
  left_join(weights_by_year_fips) %>% 
  arrange(variable_name, fips, year) %>% 
  ungroup()
head(dat, 100)
tail(dat, 100)
print(dat[25000:26000, ])

# Issue is that weighting variables are not necessary lined up with our metrics.
# Annual population from census is not included for all counties - only the ones
# with large enough population. For these, we will have to use 5-year population

# If we want to weight them, we either have to use 5-year values or assume
# population doesn't change over 5 years, or spread them out evenly between each
# 5-year known value
dat %>% 
  filter(str_detect(fips, '^50'), ) %>% 
```


## Spatial Panel Regressions

### Spatialreg

`spatialreg` for spatially weighted regressions, but not panels. First trying a single example with gini. Note that because Census data spans the change in administrative boundaries in CT, we can't use CT at all here.

```{r}
#| label: spatialreg_test
#| output: false
#| eval: false
get_str(datasets$dp_metrics_county)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Remove CT. Drop na ginis, but, get complete cases for relevant fips and years
dat <- dat %>% 
  select(fips, year, gini) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  drop_na(gini) %>% 
  complete(fips, year)
get_str(dat)

# Get our listw based on counties
lw <- neast_counties_2024 %>% 
  # filter(fips %in% dat$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Get better one for long-format
lw <- neast_counties_2024 %>% 
  right_join(dat) %>% 
  select(fips, year, gini) %>% 
  # filter(fips %in% dat$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Run regression
m1e <- errorsarlm(
  gini ~ year,
  data = dat,
  listw = lw
)
summary(m1e)
```


### Panels

#### Test Run

Trying splm package for spatially weighted panel models. This is not going well though - too much missing data means it is very difficult to get a balanced panel.

```{r}
#| label: splm
#| output: false
#| eval: false
get_str(datasets$dp_metrics_county)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Get rid of geometry, make our df a plm object
df <- dat %>% 
  select(fips, year, gini) %>%
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  na.omit() %>%
  plm::pdata.frame()
get_str(df)

# Get balanced fips
# all_years <- sort(unique(df$year))
# balanced_fips <- df %>%
#   group_by(fips) %>%
#   filter(setequal(sort(year), all_years)) %>%
#   pull(fips) %>%
#   unique()
# df <- df %>%
#   filter(fips %in% balanced_fips)

# Get our listw based on counties
lw <- neast_counties_2024 %>% 
  filter(fips %in% df$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Pooling model
pool <- spml(
  gini ~ as.numeric(year),
  data = df, 
  listw = lw,
  model = 'pooling'
)
summary(pool)

# Try FE error model with rentMedianPercHH
fe_err <- spml(
  gini ~ as.numeric(year),
  data = df, 
  listw = lw,
  model = 'within'
)
summary(fe_err)

# RE error model
re_err <- spml(
  gini ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'random'
)
summary(re_err)

# Check to use RE or FE, see whether x_it related to a_i (idiosyncratic error)
test <- sphtest(fe_err, re_err)
test
# Not sig: use RE, more efficient

# Check RE output
summary(re_err)
```


#### All Models

So the workflow for each variable is:

1. Make sure panel is balanced, i.e. see if Connecticut will be a problem. Adjust accordingly.
2. Create list weight object based on the counties that are represented by in the variable.
3. No longer bothing with FE and RE models, just doing pooling.

```{r}
#| label: spregs_over_vars
#| eval: false
get_str(datasets$dp_county_wide)
dat <- datasets$dp_county_wide %>% 
  select(fips, year, any_of(datasets$series_vars))
get_str(dat)

# Get vector of vars
vars <- names(dat)[!names(dat) %in% c('fips', 'year')]

# Map over each var and do steps 1 through 5
out <- map(vars, \(var) {
  
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var)) %>% 
      filter(str_detect(fips, '^09', negate = TRUE)) %>% 
      na.omit()
      # complete(fips, year)
      # mutate(!!sym(var) := ifelse(is.na(!!sym(var)), 0, !!sym(var)))
    
    # 1. Only use fips that produce balanced panel
    balanced_fips <- df %>%
      group_by(fips) %>%
      filter(n() == length(unique(df$year))) %>%
      pull(fips) %>%
      unique()

    df <- df %>%
      filter(fips %in% balanced_fips)

    # 2. Create lw object
    lw <- neast_counties_2024 %>%
      filter(fips %in% balanced_fips) %>%
      # filter(fips %in% df$fips) %>%
      poly2nb() %>%
      nb2listw(zero.policy = TRUE)
    
    cat('\nCreated lw')
    
    # Make df a plm object
    df <- plm::pdata.frame(df)
    
    # 3. Run pooled panel model
    formula <- as.formula(paste(var, "~ as.numeric(year)"))
    model <- spml(
      formula,
      data = df,
      listw = lw,
      index = c('fips', 'year'),
      model = 'pooling'
    )
   
    cat('\nRan model')
   
    return(list('model' = model, 'fips' = balanced_fips))
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
  
}, .progress = FALSE) %>% 
 setNames(c(vars))
out
get_str(out)
get_str(out[[1]])
summary(out$gini$model)
map(out, ~ .x['fips'])

# Save for posterity
saveRDS(out, 'data/objects/spatial_regression_outs.rds')
```

Check how they turned out:

```{r}
#| output: false
out <- readRDS('data/objects/spatial_regression_outs.rds')

# How many were successful
perc_success <- mean(map(out, class) == 'list') * 100

# Check out errors only
errors <- out %>% 
  keep(is.character)
errors
error_vars <- names(errors)
```

`{r} length(error_vars)` models failed: `{r} paste(error_vars, collapse = ', ')`. This includes all the Forest Inventory Analysis vars. Note to check this and see if they are missing data for lots of counties.


### Results

Check out the successful models:

```{r}
#| label: spatial_panel_table
model_outs <- out %>% 
  keep(is.list)
# get_str(model_outs)
# get_str(model_outs[[1]])
# get_str(model_outs[[1]][[1]])

# Check model type
# model_outs[[1]][[1]]$type
# table(map_chr(model_outs, ~ .x[[1]]$type))
# They are all RE models - no FE


## Test out outcomes
# sum <- summary(model_outs[[1]][[1]])
# coefs <- sum$CoefTable[2, ]

# Make a table
spatial_table <- imap(model_outs, ~ {
  tryCatch({
    
    # Pull model output results and coefficients
    sum <- summary(.x[[1]])
    coefs <- sum$CoefTable[2, ]
    
    df <- data.frame(
      n = length(.x$fips),
      # type = .x[[1]]$type,
      var = .y,
      # phi = sum$errcomp[1],
      rho = sum$ErrCompTable[[1]],
      est = coefs[1],
      se = coefs[2],
      t = coefs[3],
      p = coefs[4]
    ) 
    
    # Figure out good or bad outcome
    # If sig, then check reverse. else blank
    # If reverse, reverse it, else same
    # If up, good, else bad
    df <- df %>% 
      mutate(
        sig = ifelse(p < 0.05, TRUE, FALSE),
        posneg = case_when(
          est > 0 ~ 'pos',
          est < 0 ~ 'neg',
          .default = ''
        ),
        reverse = ifelse(.y %in% reverse, TRUE, FALSE),
        direction = case_when(
          sig == TRUE & reverse == TRUE & posneg == 'pos' ~ 'neg',
          sig == TRUE & reverse == TRUE & posneg == 'neg' ~ 'pos',
          sig == TRUE & reverse == FALSE ~ as.character(posneg),
          .default = ''
        ),
        outcome = case_when(
          sig == FALSE ~ '',
          sig == TRUE & direction == 'pos' ~ 'better',
          sig == TRUE & direction == 'neg' ~ 'worse',
          .default = '-'
        )
      ) %>% 
      select(-c(posneg, direction))
    return(df)
  },
    error = function(e) {
      return(paste('Error:', e))
    }
  )
}) %>% 
  bind_rows()
get_str(spatial_table)

# Round off and make table
spatial_table %>% 
  mutate(
    across(c(rho:p), ~ format(round(.x, 3), nsmall = 3))
    # type = case_when(
    #   str_detect(type, 'random') ~ 'RE',
    #   .default = NA
    # )
  ) %>% 
  get_reactable(
    defaultColDef = colDef(minWidth = 50),
    columns = list(
      n = colDef(minWidth = 25),
      var = colDef(minWidth = 200),
      est = colDef(
        minWidth = 100,
        name = 'β'
      ),
      outcome = colDef(minWidth = 75),
      # rho = colDef(name = 'ρ'),
      # phi = colDef(name = 'φ'),
      p = colDef(name = 'p-value'),
      'p-value' = colDef(minWidth = 100)
    )
  )
```

- n: Number of counties included in regression
<!-- - type: FE = fixed effects (every county has $\beta$), RE = random effects (counties have different $\beta$s) -->
- var: Metric
<!-- - $\phi$: Spatial autocorrelation in idiosyncratic attributes (counties) -->
- $\rho$: Spatial autocorrelation in residual errors (omitted variables)
- $\beta$: Regression coefficient of year on the variable
- se: standard error of $\beta$
- t: test statistic
- p-value: p-value
- sig: whether p is < 0.05
- reverse: whether the metric has been reversed so that lower numbers are better and higher numbers are worse (like food insecurity)
- outcome: if the regression is significant, is it getting better or worse? Accounts for the 'reverse' variable