---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  conflicted,
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  caret,
  knitr,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  spatialreg,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr,
  flextable,
  officer,
  lmtest,
  sandwich
)

pacman::p_load_gh(
  'ChrisDonovan307/projecter', 
  'Food-Systems-Research-Institute/SMdata'
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

results <- list()
datasets <- readRDS('data/data_paper/datasets.rds')
```

## Linear Models

First trying linear models with time as independent variable. Normalizing dataset before we do anything else.

```{r}
#| label: normalize
#| output: false
get_str(datasets$dp_metrics_county)

# Function to min max then scale to 0 to 100 for percent change
min_max <- function(x) {
  normed <- (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  normed * 100
}

vars <- unique(datasets$dp_metrics_county$variable_name)
datasets$dp_metrics_county <- map(vars, ~ {
  datasets$dp_metrics_county %>%
    filter(variable_name == .x) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
get_str(datasets$dp_metrics_county)  
```


### Regressions

First create a weights crosswalk that links `variable_name`s to weighting variables.

```{r}
#| label: weights_crosswalk
#| output: false
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Population (5 year, smooth)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres operated, smooth',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5Year',
    'population5YearSmooth',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'acresOperatedSmooth',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)
weights_crosswalk

# Link weight vars to the appropriate variables
weights_crosswalk <- data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric) %>% 
  na.omit()
get_str(weights_crosswalk)
weights_crosswalk
```

Run regressions for all time series variables where time points > 1.

```{r}
#| label: regressions
#| output: false
#| eval: false
# Get a crosswalk to add state as a fixed effect
state_link <- fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

res <- map(datasets$series_vars, ~ {
  tryCatch(
    {
      df <- datasets$dp_metrics_county %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year),
          state_fips = str_sub(fips, end = 2)
        ) %>% 
        left_join(state_link, by = join_by(state_fips == fips))
      
      # Add weighting variable
      weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == .x]
      weight_df <- datasets$dp_weights
      if (!is.na(weight_var) && length(weight_var) > 0) {
        weight_df <- weight_df %>% 
          filter(variable_name == weight_var)
      }
      weight_df <- weight_df %>% 
        mutate(
          value = as.numeric(value), 
          year = as.numeric(year)
        ) %>% 
        rename(
          weight_value = value, 
          weight_name = variable_name,
          weight_year = year
        )
      # get_str(weight_df)
      
      # Depending on whether weight has a year, do appropriate join
      # Note that we might be losing some years here if weighting var doesn't
      # cover the variable enough
      if (all(is.na(weight_df$weight_year))) {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips))
      } else {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips, year == weight_year))
      }
      
      # Reduce to county fips only. Remove dupes 
      df <- df %>%
        filter(str_length(fips) == 5) %>% 
        distinct()
      
      # Weighted OLS 
      model <- lm(
        normed_value ~ year + state_code, 
        data = df,
        weights = weight_value
      )
      return(model)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(datasets$series_vars)) %>%
  discard(\(x) is.null(x))
get_str(res, 1)
get_str(res)

# Robust SEs
pacman::p_load(lmtest)
coeftest(res[[1]])

# Save immediate output
saveRDS(res, 'data/objects/weighted_ols_outs.rds')
```


### Residuals

Check residuals of outputs:

```{r}
#| label: lm_residuals
#| freeze: true
#| warnings: false
#| outputs: false
res <- readRDS('data/objects/weighted_ols_outs.rds')
res_plots <- map(res, ~ {
  try(performance::check_model(.x))
}) %>% 
  purrr::keep(~ length(.x) > 1)
# get_str(res_plots)
saveRDS(res_plots, 'data/objects/weighted_ols_residuals.rds')
```


```{r}
#| fig-width: 10
#| fig-height: 10
res_plots <- readRDS('data/objects/weighted_ols_residuals.rds')
res_plots
```

Now that we have regression outputs, let's put them in a table to see which ones change significantly over time.

```{r}
#| label: make_reg_table
#| output: false
# Put together one table of results from all models, add stars
res <- readRDS('data/objects/weighted_ols_outs.rds')
get_str(res)

# Check for NA outputs
res[[1]]$coefficients[['year']]
coefs <- map(res, ~ .x$coefficients[['year']])
names(coefs[is.na(coefs)])

# Mobility disability is NA because there is only 1 year after we join with 
# weighting variable. So we need to just throw it out if we can't get smoothed
# population for 2024.
res <- res %>% 
  discard(~ is.na(.x$coefficients[[2]]))

# test
vcov = vcovHC(res[[1]], type = 'HC0')
test <- coeftest(res[[1]], vcov = vcov)
test['year', ]
cis <- coefci(res[[1]], vcov. = vcov, level = 0.95)
cis['year', ]

res_df <- imap(res, ~ {
  vcov = vcovHC(.x, type = 'HC0')
  df <- coeftest(.x, vcov = vcov) %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
  conf_ints <- coefci(.x)['year', ]
  df$ci_low <- conf_ints[1]
  df$ci_high <- conf_ints[2]
  df
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many time series metrics total
(n_metrics <- nrow(res_df))

# How many change over time significantly
(
  n_change <- res_df %>% 
    filter(sig == '*') %>% 
    nrow()
)

# Percentage significant
(perc_sig <- mean(res_df$sig == '*', na.rm = TRUE) * 100)
```

Note that we only have 2 years for the mobility disability metric - 2023 and 2024. But we don't have a smoothed 5-year population for 2024, only 2023. So we can only use a single year, which means there is no variation in year. So we are dropping this from the analysis.

Out of `{r} n_metrics` metrics with > 1 time point, we have `{r} n_change` metrics that change significantly over time (`{r} round(perc_sig, 2)`%).

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch.

```{r}
#| label: make_sense_of_trends
#| output: false
# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
raw_outputs <- res_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(data_paper_meta)
directions <- data_paper_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics
model_outputs <- raw_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# See what we lost
(lost <- setdiff(raw_outputs$term, model_outputs$term))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).

Check out table with outputs for each regression. Filter by significance and outcome. Trend is just whether it is going up or down, outcome is whether that is good or bad.


### Table

```{r}
#| label: reg_table
## Add framework info
framed_outputs <- data_paper_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(model_outputs, by = join_by(variable_name == term))


## Save some outputs for paper
# Raw csv for reference
write_csv(
  framed_outputs,
  'outputs/lm_reg_outputs.csv'
)

# Make table for Word
framed_outputs %>% 
  select(-c(
    dimension,
    index, 
    indicator,
    variable_name,
    trend,
    reverse,
    sig,
    starts_with('ci')
  )) %>% 
  flextable() %>% 
  flextable::font(part = "all", fontname = "Times New Roman") %>% 
  save_as_docx(
    path = 'outputs/lm_reg_outputs.docx'
  )


## Create interactive table for Quarto
get_reactable(
  framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```


### Figure

Figure below shows 95% confidence intervals for the coefficients of each metric, including state as a covariate.

```{r}
#| label: regression_figure
# get_str(framed_outputs)
lm_figure <- framed_outputs %>% 
  mutate(
    across(c(estimate, starts_with('ci')), as.numeric),
    outcome = ifelse(is.na(outcome), 'no change', outcome) %>% 
      snakecase::to_title_case()
  ) %>% 
  filter(!variable_name %in% c('voterTurnout', 'mobilityDisability')) %>% 
  ggplot(aes(x = estimate, y = reorder(metric, estimate))) +
  geom_point(aes(color = outcome)) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
  geom_vline(xintercept = 0) +
  scale_color_manual(values = c('darkgreen', 'grey', 'purple')) +
  theme_minimal() +
  labs(
    x = 'Average Relative Percent Change per Year',
    y = 'Metric',
    color = 'Outcome'
  )
ggsave(
  'outputs/lm_conf_ints.png',
  plot = lm_figure,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)
lm_figure
```

```{r}
#| label: get_weights
#| eval: false
#| include: false

### Weights section
# This did not work out well. We use the crosswalk above but we are ditching 
# the idea of trying to join them all at once. Instead, we make joins with
# weights for each specific variable in the regression process.

# Adding weight variables eventually
get_str(datasets$dp_weights)
datasets$dp_weights$variable_name %>% 
  unique()
get_str(data_paper_meta)
data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name)

# Define weights crosswalk from metrics to variable names here
# TODO: Work this into data export from SMdata package
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Population (5 year, smooth)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5Year',
    'population5YearSmooth',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)
weights_crosswalk

# Take our county data and join weight variable names to it
weights_crosswalk <- data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric) %>% 
  na.omit()
get_str(weights_crosswalk)
weights_crosswalk
```


## Spatial Panel Regressions

### Spatialreg

`spatialreg` for spatially weighted regressions, but not panels. First trying a single example with gini. Note that because Census data spans the change in administrative boundaries in CT, we can't use CT at all here.

```{r}
#| label: spatialreg_test
#| output: false
#| eval: false
get_str(datasets$dp_metrics_county)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Remove CT. Drop na ginis, but, get complete cases for relevant fips and years
dat <- dat %>% 
  select(fips, year, gini) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  drop_na(gini) %>% 
  complete(fips, year)
get_str(dat)

# List weights
lw <- neast_counties_2024 %>% 
  right_join(dat) %>% 
  select(fips, year, gini) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Run regression
m1e <- errorsarlm(
  gini ~ year,
  data = dat,
  listw = lw
)
summary(m1e)
```


#### All Models

Use `errorsarlm` on all metrics.

```{r}
#| label: spatial_reg_all
#| eval: false
# herehere
get_str(datasets$dp_metrics_county)
state_link <- fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  ) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>% 
  mutate(link = str_sub(fips, end = 2)) %>% 
  left_join(state_link, by = join_by(link == fips)) %>% 
  select(-link)
get_str(dat)

# Get vector of vars
(vars <- names(dat)[!names(dat) %in% c('fips', 'year')])

tic()
get_time()
plan(multisession, workers = availableCores(omit = 1))
out <- future_map(vars, \(var) {
# out <- map(vars[1:5], \(var) {
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var), state_code) %>% 
      drop_na(!!sym(var))
 
    # List weights
    lw <- neast_counties_2024 %>% 
      right_join(df) %>% 
      select(fips, year, !!sym(var)) %>% 
      poly2nb() %>% 
      nb2listw(zero.policy = TRUE)   
    cat('\nCreated lw')

    # Run error sar model
    formula <- as.formula(paste(var, "~ as.numeric(year) + state_code"))
    model <- errorsarlm(
      formula,
      data = df,
      listw = lw
    ) 
    cat('\nRan model')
   
    return(model)
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
}, .progress = FALSE) %>% 
 setNames(c(vars))
plan(sequential)
toc()

out

# get_str(out)
# get_str(out[[1]])
# summary(out[[1]]$model)
# map(out, ~ .x['fips'])
# map(out, ~ summary(.x[[1]]))

saveRDS(out, 'data/objects/errorsarlm_outs.rds')
```

```{r}
sar_outs <- readRDS('data/objects/errorsarlm_outs.rds')
get_str(sar_outs)
```


### Panels

#### Test Run

Trying splm package for spatially weighted panel models. This is not going well though - too much missing data means it is very difficult to get a balanced panel.

```{r}
#| label: splm_test_run
#| output: false
#| eval: false
get_str(datasets$dp_metrics_county)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Get rid of geometry, make our df a plm object
df <- dat %>% 
  select(fips, year, gini) %>%
  # select(fips, year, otyAnnualAvgEstabsCountPctChgNAICS11) %>%
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  na.omit() %>%
  plm::pdata.frame()
get_str(df)

df %>% 
  group_by(year, fips) %>% 
  summarize(n()) %>% 
  arrange(year, fips) %>% 
  print(n = 500)

# Get balanced fips
all_years <- sort(unique(df$year))
balanced_fips <- df %>%
  group_by(fips) %>%
  filter(setequal(sort(year), all_years)) %>%
  pull(fips) %>%
  unique()
df <- df %>%
  filter(fips %in% balanced_fips)

# Get our listw based on counties
lw <- neast_counties_2024 %>% 
  filter(fips %in% df$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Pooling model
pool <- spml(
  gini ~ as.numeric(year),
  data = df, 
  listw = lw,
  model = 'pooling'
)
summary(pool)

# Try FE error model
fe_err <- spml(
  gini ~ as.numeric(year),
  data = df, 
  listw = lw,
  model = 'within'
)
summary(fe_err)

# RE error model
re_err <- spml(
  gini ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'random'
)
summary(re_err)

# Try lagsarlm
lagsar <- lagsarlm(
  # otyAnnualAvgEstabsCountPctChgNAICS11 ~ as.numeric(year),
  gini ~ as.numeric(year),
  data = df,
  listw = lw
)
summary(lagsar)

# # Check to use RE or FE, see whether x_it related to a_i (idiosyncratic error)
# test <- sphtest(fe_err, re_err)
# test
# # Not sig: use RE, more efficient
# 
# # Check RE output
# summary(re_err)
```


#### All Models

```{r}
#| label: spml_models
#| eval: false
get_str(datasets$dp_county_wide)
dat <- datasets$dp_county_wide %>% 
  select(fips, year, any_of(datasets$series_vars))
get_str(dat)

# Get vector of vars
vars <- names(dat)[!names(dat) %in% c('fips', 'year')]

# Map over each var and do steps 1 through 5
out <- map(vars, \(var) {
  
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var)) %>% 
      filter(str_detect(fips, '^09', negate = TRUE)) %>% 
      na.omit()
      # complete(fips, year)
      # mutate(!!sym(var) := ifelse(is.na(!!sym(var)), 0, !!sym(var)))
    
    # 1. Only use fips that produce balanced panel
    balanced_fips <- df %>%
      group_by(fips) %>%
      filter(n() == length(unique(df$year))) %>%
      pull(fips) %>%
      unique()

    df <- df %>%
      filter(fips %in% balanced_fips)

    # 2. Create lw object
    lw <- neast_counties_2024 %>%
      filter(fips %in% balanced_fips) %>%
      # filter(fips %in% df$fips) %>%
      poly2nb() %>%
      nb2listw(zero.policy = TRUE)
    
    cat('\nCreated lw')
    
    # Make df a plm object
    df <- plm::pdata.frame(df)
    
    # 3. Run pooled panel model
    formula <- as.formula(paste(var, "~ as.numeric(year)"))
    model <- spml(
      formula,
      data = df,
      listw = lw,
      index = c('fips', 'year'),
      model = 'pooling',
      spatial.error = 'b'
      # lag = TRUE
    )
   
    cat('\nRan model')
   
    return(list('model' = model, 'fips' = balanced_fips))
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
  
}, .progress = FALSE) %>% 
 setNames(c(vars))
out
get_str(out)
get_str(out[[1]])
summary(out[[1]]$model)
map(out, ~ .x['fips'])
map(out, ~ summary(.x[[1]]))

# Save for posterity
saveRDS(out, 'data/objects/spatial_regression_outs.rds')
```


#### All Models Weighted

```{r}
#| label: spml_models_weighted
#| eval: false
# TODO: fix this. we broke it here somewhere.
get_str(datasets$dp_county_wide)
dat <- datasets$dp_county_wide %>% 
  select(fips, year, any_of(datasets$series_vars))
get_str(dat)

# Get vector of vars
vars <- names(dat)[!names(dat) %in% c('fips', 'year')]


## Normalization
get_str(dat)
dat <- dat %>% 
  mutate(across(all_of(vars), ~ min_max(.x)))
get_str(dat)  
map(dat, ~ range(.x, na.rm = TRUE))

 
## Regressions
out <- map(vars, \(var) {
# out <- map('treeDiversity', \(var) {
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var)) %>% 
      filter(str_detect(fips, '^09', negate = TRUE)) %>% 
      na.omit()
    
    ## Add weights
    weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == var]
    weight_df <- datasets$dp_weights
    if (!is.na(weight_var) && length(weight_var) > 0) {
      weight_df <- weight_df %>%
        filter(variable_name == weight_var)
    }
    weight_df <- weight_df %>%
      mutate(
        value = as.numeric(value),
        year = as.numeric(year)
      ) %>%
      rename(
        weight_value = value,
        weight_name = variable_name,
        weight_year = year
      )
    # Do join depending on whether weight var has a year
    if (all(is.na(weight_df$weight_year))) {
      df <- df %>%
        inner_join(weight_df, by = join_by(fips))
    } else {
      df <- df %>%
        inner_join(weight_df, by = join_by(fips, year == weight_year))
    }
    # Weight the variable appropriately
    var_weighted <- paste0(var, '_weighted')
    df <- df %>%
      mutate(!!sym(var_weighted) := !!sym(var) * sqrt(weight_value)) %>% 
      drop_na(!!sym(var_weighted))
    
    ##  Only use fips that produce balanced panel
    balanced_fips <- df %>%
      group_by(fips) %>%
      filter(n() == length(unique(df$year))) %>%
      pull(fips) %>%
      unique()
    df <- df %>%
      filter(fips %in% balanced_fips)
    
    ## Create lw object
    lw <- neast_counties_2024 %>%
      filter(fips %in% balanced_fips) %>%
      filter(fips %in% df$fips) %>%
      poly2nb() %>%
      nb2listw(zero.policy = TRUE)
    cat('\nCreated lw')
    
    ## Run pooled panel model
    df <- plm::pdata.frame(df)
    # formula <- as.formula(paste(var_weighted, "~ as.numeric(year)"))
    formula <- as.formula(paste(var_weighted, "~ as.numeric(year)"))
    model <- spml(
      formula,
      data = df,
      listw = lw,
      index = c('fips', 'year'),
      model = 'pooling'
    )
   
    cat('\nRan model')
   
    return(list('model' = model, 'fips' = balanced_fips))
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
}, .progress = FALSE) %>% 
 setNames(c(vars))
out
out[[2]]
out$cropDiversity # check for smoothed weighting var
get_str(out)
get_str(out[[1]])
summary(out$gini$model)
map(out, ~ .x['fips'])
map(out, ~ summary(.x[[1]]))
# Getting errors because there are NO balanced fips

# Save for posterity
saveRDS(out, 'data/objects/spatial_panel_outs.rds')
```

Check how they turned out:

```{r}
#| output: false
out <- readRDS('data/objects/spatial_panel_outs.rds')
get_str(out)

# How many were successful
(perc_success <- mean(map(out, class) == 'list') * 100)

# Check out errors only
errors <- out %>% 
  keep(is.character)
errors
(error_vars <- names(errors))
```

`{r} length(error_vars)` models failed: `{r} paste(error_vars, collapse = ', ')`.

#### Results

Check out the successful models:

##### Table

```{r}
#| label: spatial_panel_table
model_outs <- out %>% 
  keep(is.list)

# Make a table
spatial_table <- imap(model_outs, ~ {
  tryCatch({
    
    # Pull model output results and coefficients
    sum <- summary(.x[[1]])
    coefs <- sum$CoefTable[2, ]
    
    # Add confidence intervals
    conf_ints <- confint(.x[[1]])[2, ] %>% 
      as.vector()
    # df$ci_low <- conf_ints[1]
    # df$ci_high <- conf_ints[2]
  
    df <- data.frame(
      n = length(.x$fips),
      # type = .x[[1]]$type,
      var = .y,
      # phi = sum$errcomp[1],
      rho = sum$ErrCompTable[[1]],
      est = coefs[1],
      se = coefs[2],
      t = coefs[3],
      ci_low = conf_ints[1],
      ci_high = conf_ints[2],
      p = coefs[4]
    ) 
    
    # Figure out good or bad outcome
    # If sig, then check reverse. else blank
    # If reverse, reverse it, else same
    # If up, good, else bad
    df <- df %>% 
      mutate(
        sig = ifelse(p < 0.05, TRUE, FALSE),
        posneg = case_when(
          est > 0 ~ 'pos',
          est < 0 ~ 'neg',
          .default = ''
        ),
        reverse = ifelse(.y %in% reverse, TRUE, FALSE),
        direction = case_when(
          sig == TRUE & reverse == TRUE & posneg == 'pos' ~ 'neg',
          sig == TRUE & reverse == TRUE & posneg == 'neg' ~ 'pos',
          sig == TRUE & reverse == FALSE ~ as.character(posneg),
          .default = ''
        ),
        outcome = case_when(
          sig == FALSE ~ '',
          sig == TRUE & direction == 'pos' ~ 'better',
          sig == TRUE & direction == 'neg' ~ 'worse',
          .default = '-'
        )
      ) %>% 
      select(-c(posneg, direction))
    return(df)
  },
    error = function(e) {
      return(paste('Error:', e))
    }
  )
}) %>% 
  bind_rows()
# get_str(spatial_table)


## Make table for Word
spatial_table %>% 
  select(n:t, p, outcome) %>% 
  mutate(across(c(rho:p), ~ format(round(.x, 3), nsmall = 3))) %>% 
  flextable() %>% 
  flextable::font(part = "all", fontname = "Times New Roman") %>% 
  save_as_docx(
    path = 'outputs/spatial_reg_outputs.docx'
  )


## Round off and make table for Quarto
spatial_table %>% 
  mutate(
    across(c(rho:p), ~ format(round(.x, 3), nsmall = 3))
  ) %>% 
  get_reactable(
    defaultColDef = colDef(minWidth = 50),
    columns = list(
      n = colDef(minWidth = 50),
      var = colDef(minWidth = 200),
      est = colDef(
        minWidth = 100,
        name = 'β'
      ),
      outcome = colDef(minWidth = 75),
      rho = colDef(name = 'ρ'),
      p = colDef(name = 'p-value'),
      'p-value' = colDef(minWidth = 100)
    )
  )
```

Issue here is that we have to drop many counties for some variables. The `n` column shows number of counties included. Some are well below half - not sure this is a great way to assess change if that is the case.

- n: Number of counties included in regression
<!-- - type: FE = fixed effects (every county has $\beta$), RE = random effects (counties have different $\beta$s) -->
- var: Metric
<!-- - $\phi$: Spatial autocorrelation in idiosyncratic attributes (counties) -->
- $\rho$: Spatial autocorrelation in residual errors (omitted variables)
- $\beta$: Regression coefficient of year on the variable
- se: standard error of $\beta$
- t: test statistic
- p-value: p-value
- sig: whether p is < 0.05
- reverse: whether the metric has been reversed so that lower numbers are better and higher numbers are worse (like food insecurity)
- outcome: if the regression is significant, is it getting better or worse? Accounts for the 'reverse' variable

##### Figure

```{r}
#| label: spatial_reg_figure
# get_str(spatial_framed_outputs)
# get_str(spatial_table)
spatial_framed_outputs <- data_paper_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(spatial_table, by = join_by(variable_name == var))
# get_str(spatial_framed_outputs)

# Save some outputs to OneDrive
write_csv(
  spatial_framed_outputs,
  'outputs/spatial_reg_outputs.csv'
)

flex <- spatial_framed_outputs %>% 
  select(-variable_name) %>% 
  as_flextable(show_coltype = FALSE)
flextable::save_as_docx(
  flex,
  path = 'outputs/spatial_reg_outputs.docx'
)
out <- spatial_framed_outputs %>% 
  mutate(
    across(c(est, starts_with('ci')), as.numeric),
    outcome = case_when(
      outcome == '' ~ 'no change',
      .default = outcome
    ) %>% 
      snakecase::to_title_case()
  ) %>% 
  ggplot(aes(x = est, y = reorder(metric, est))) +
  geom_point(aes(color = outcome)) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
  geom_vline(xintercept = 0) +
  scale_color_manual(values = c('darkgreen', 'grey', 'purple')) +
  theme_minimal() +
  labs(
    x = 'Average Relative Weighted Percent Change per Year',
    y = 'Metric',
    color = 'Outcome'
  )
ggsave(
  'outputs/spatial_reg_conf_ints.png',
  plot = out,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)
out
```
