---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  caret,
  knitr,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')

# Load SMdata package from repos folder
sm_load()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

results <- list()
```

## Linear Models

First trying linear models with time as independent variable. Not really using these for anything but comparison though. Normalizing dataset before we do anything else.

```{r}
#| label: normalize
#| output: false
datasets <- readRDS('data/data_paper/datasets.rds')
get_str(datasets$dp_metrics_county)

# Function to min max 0 to 1
min_max <- function(x) {
  normed <- (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  normed * 100
}

vars <- unique(datasets$dp_metrics_county$variable_name)
datasets$dp_metrics_county <- map(vars, ~ {
  datasets$dp_metrics_county %>%
    filter(variable_name == .x) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
get_str(datasets$dp_metrics_county)  
```

```{r}
#| eval: false
#| echo: false
# Adding weight variables eventually
get_str(datasets$dp_weights)
datasets$dp_weights$variable_name %>% 
  unique()
get_str(data_paper_meta)
data_paper_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% datasets$dp_metrics_county$variable_name)
```

Run regressions:

```{r}
#| label: trend_analysis
#| output: false
# Get a crosswalk to add state as a fixed effect
state_link <- fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

res <- map(datasets$series_vars, ~ {
  tryCatch(
    {
      df <- datasets$dp_metrics_county %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year),
          state_fips = str_sub(fips, end = 2)
        ) %>% 
        left_join(state_link, by = join_by(state_fips == fips))
      model <- lm(normed_value ~ year + state_code, data = df)
      return(model)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(datasets$series_vars)) %>%
  discard(\(x) is.null(x))
```

Now that we have regression outputs, let's put them in a table to see which ones were significant. 

```{r}
#| label: make_table
#| output: false
# Put together one table of results from all models, add stars
res_df <- imap(res, ~ {
  df <- .x %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
  conf_ints <- confint(.x)['year', ] %>% 
    as.vector()
  df$ci_low <- conf_ints[1]
  df$ci_high <- conf_ints[2]
  df
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many metrics
n_metrics <- nrow(res_df)

# How many are significant
perc_sig <- mean(res_df$sig == '*') * 100
```

We have `{r} n_metrics` metrics, `{r} round(perc_sig, 1)` of which vary significantly over time.

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch.

```{r}
#| label: make_sense_of_trends
#| output: false
# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
raw_outputs <- res_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(data_paper_meta)
directions <- data_paper_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose 4 metrics
model_outputs <- raw_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# See what we lost
lost <- setdiff(raw_outputs$term, model_outputs$term) %>% 
  paste0(collapse = ', ')

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
(percs <- prop.table(tab) * 100)
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} lost`).

Check out table with outputs for each regression. Filter by significance and outcome. Trend is just whether it is going up or down, outcome is whether that is good or bad.


### Table

```{r}
#| label: model_table
# Add framework info
framed_outputs <- data_paper_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(model_outputs, by = join_by(variable_name == term))

write_csv(
  framed_outputs,
  'outputs/lm_reg_outputs.csv'
)

get_reactable(
  framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```


### Figure

```{r}
get_str(framed_outputs)

out <- framed_outputs %>% 
  mutate(
    across(c(estimate, starts_with('ci')), as.numeric),
    outcome = ifelse(is.na(outcome), 'no change', outcome) %>% 
      snakecase::to_title_case()
  ) %>% 
  ggplot(aes(x = estimate, y = reorder(metric, estimate))) +
  geom_point(aes(color = outcome)) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
  geom_vline(xintercept = 0) +
  scale_color_manual(values = c('darkgreen', 'grey', 'purple')) +
  theme_minimal() +
  labs(
    x = 'Average Percent Change per Year',
    y = 'Metric',
    color = 'Outcome'
  )
ggsave(
  'outputs/lm_conf_ints.png',
  plot = out,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)
out
```


## Spatial Panel Regressions

### Setup

```{r}
#| label: splm
#| output: false
#| eval: false
get_str(datasets$dp_metrics_county)

dat <- datasets$dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Get rid of geometry, make our df a plm object
df <- dat %>% 
  # select(fips, year, gini) %>% 
  select(fips, year, otyAnnualAvgEstabsCountPctChgNAICS11) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  na.omit() %>%
  plm::pdata.frame()
get_str(df)

# Get balanced fips
all_years <- sort(unique(df$year))
balanced_fips <- df %>%
  group_by(fips) %>%
  filter(setequal(sort(year), all_years)) %>%
  pull(fips) %>%
  unique()
df <- df %>%
  filter(fips %in% balanced_fips)

# Get our listw based on counties
lw <- neast_counties_2024 %>% 
  filter(fips %in% df$fips) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Try FE error model with rentMedianPercHH
fe_err <- spml(
  # gini ~ as.numeric(year), 
  otyAnnualAvgEstabsCountPctChgNAICS11 ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'within'
)
summary(fe_err)

# RE error model
re_err <- spml(
  gini ~ as.numeric(year), 
  data = df, 
  listw = lw,
  model = 'random'
)
summary(re_err)

# Check to use RE or FE, see whether x_it related to a_i (idiosyncratic error)
test <- sphtest(fe_err, re_err)
test
# Not sig: use RE, more efficient

# Check RE output
summary(re_err)
```


### Run Models

So the workflow for each variable is:

1. Make sure panel is balanced, i.e. see if Connecticut will be a problem. Adjust accordingly.
2. Create list weight object based on the counties that are represented by in the variable. I think we have to do this for each variable individually... should double check though.
3. Run FE and RE model with year as numeric.
4. Hausman test with FE and RE model. If sig, report FE, if not, report RE

Try it:

```{r}
#| label: spregs_over_vars
#| eval: false
get_str(dat)

# Get vector of vars
vars <- names(dat)[!names(dat) %in% c('fips', 'year')]

# Map over each var and do steps 1 through 5
# plan(multisession, cores = availableCores(omit = 2))
out <- map(vars[2], \(var) {
  
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var)) %>% 
      na.omit()
    
    # 1. Only use fips that produce balanced panel
    balanced_fips <- df %>% 
      group_by(fips) %>% 
      filter(n() == length(unique(df$year))) %>% 
      pull(fips) %>% 
      unique()
    
    df <- df %>% 
      filter(fips %in% balanced_fips)
    
    # 2. Create lw object
    # lw <- neast_counties_2024 %>% 
    #   filter(fips %in% balanced_fips) %>% 
    #   filter(fips %in% df$fips) %>%
    #   poly2nb() %>% 
    #   nb2listw(zero.policy = TRUE)
    
    # Keep only relevant counties
    county_sf <- neast_counties_2024 %>%
      filter(fips %in% balanced_fips)
    
    # Ensure same order as df
    county_sf <- county_sf %>%
      arrange(match(fips, unique(df$fips)))
    
    # Neighbor list
    nb <- poly2nb(county_sf)
    if (any(card(nb) == 0)) {
      stop("One or more counties have no neighbors!")
    }
    
    # Create listw with correct region.id
    lw <- nb2listw(nb, style = "W", zero.policy = TRUE)
    # attr(lw, "region.id") <- as.character(county_sf$fips)
    stopifnot(length(attr(lw, "region.id")) == length(unique(df$fips)))
    stopifnot(all(attr(lw, "region.id") %in% rownames(df)))
    ##
    cat('\nCreated lw')

    
    
    
    # Make df a plm object
    df <- plm::pdata.frame(df)
    
    # 3. Run FE and RE
    formula <- as.formula(paste(var, "~ as.numeric(year)"))
    fe <- spml(
      formula,
      data = df, 
      listw = lw,
      model = 'within'
    )
    re <- spml(
      formula,
      data = df, 
      listw = lw,
      model = 'random'
    )
    cat('\nRan models')
    
    # 4. Hausman test. If sig, use FE
    test <- sphtest(fe, re)
    if (test$p.value < 0.05) {
      model <- fe
    } else {
      model <- re
    }
    cat('\nRan Hausman')
    return(list('model' = model, 'fips' = balanced_fips))
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
  
}, .progress = FALSE) %>% 
 setNames(c(vars[2]))
# plan(sequential)
out
get_str(out)

# Save for posterity
saveRDS(out, 'data/objects/spatial_regression_outs.rds')
```

<!-- Check how they turned out: -->

```{r}
#| eval: false
out <- readRDS('data/objects/spatial_regression_outs.rds')

# How many were successful
perc_success <- mean(map(out, class) == 'list') * 100

get_str(out[[1]])
get_str(out)
out

# Check out errors only
errors <- out %>% 
  keep(is.character)
errors
error_vars <- names(errors)
```

### Results

Check out the successful models:

```{r}
#| label: spatial_panel_table
#| eval: false
model_outs <- out %>% 
  keep(is.list)
# get_str(model_outs)
# get_str(model_outs[[1]])
# get_str(model_outs[[1]][[1]])

# Check model type
# model_outs[[1]][[1]]$type
# table(map_chr(model_outs, ~ .x[[1]]$type))
# They are all RE models - no FE


## Test out outcomes
# sum <- summary(model_outs[[1]][[1]])
# coefs <- sum$CoefTable[2, ]

# Make a table
spatial_table <- imap(model_outs, ~ {
  tryCatch({
    
    # Pull model output results and coefficients
    sum <- summary(.x[[1]])
    coefs <- sum$CoefTable[2, ]
    
    df <- data.frame(
      n = length(.x$fips),
      type = .x[[1]]$type,
      var = .y,
      phi = sum$errcomp[1],
      rho = sum$errcomp[2],
      est = coefs[1],
      se = coefs[2],
      t = coefs[3],
      p = coefs[4]
    ) 
    
    # Figure out good or bad outcome
    # If sig, then check reverse. else blank
    # If reverse, reverse it, else same
    # If up, good, else bad
    df <- df %>% 
      mutate(
        sig = ifelse(p < 0.05, TRUE, FALSE),
        posneg = case_when(
          est > 0 ~ 'pos',
          est < 0 ~ 'neg',
          .default = ''
        ),
        reverse = ifelse(.y %in% reverse, TRUE, FALSE),
        direction = case_when(
          sig == TRUE & reverse == TRUE & posneg == 'pos' ~ 'neg',
          sig == TRUE & reverse == TRUE & posneg == 'neg' ~ 'pos',
          sig == TRUE & reverse == FALSE ~ as.character(posneg),
          .default = ''
        ),
        outcome = case_when(
          sig == FALSE ~ '',
          sig == TRUE & direction == 'pos' ~ 'better',
          sig == TRUE & direction == 'neg' ~ 'worse',
          .default = '-'
        )
      ) %>% 
      select(-c(posneg, direction))
    return(df)
  },
    error = function(e) {
      return(paste('Error:', e))
    }
  )
}) %>% 
  bind_rows()
# get_str(spatial_table)

# Round off and make table
spatial_table %>% 
  mutate(
    across(c(phi:p), ~ format(round(.x, 3), nsmall = 3)),
    type = case_when(
      str_detect(type, 'random') ~ 'RE',
      .default = NA
    )
  ) %>% 
  get_reactable(
    defaultColDef = colDef(minWidth = 50),
    columns = list(
      n = colDef(minWidth = 25),
      var = colDef(minWidth = 200),
      est = colDef(
        minWidth = 100,
        name = 'β'
      ),
      outcome = colDef(minWidth = 75),
      rho = colDef(name = 'ρ'),
      phi = colDef(name = 'φ'),
      p = colDef(name = 'p-value'),
      'p-value' = colDef(minWidth = 100)
    )
  )
```

<!-- - n: Number of counties included in regression -->
<!-- - type: FE = fixed effects (every county has $\beta$), RE = random effects (counties have different $\beta$s) -->
<!-- - var: Metric -->
<!-- - $\phi$: Spatial autocorrelation in idiosyncratic attributes (counties) -->
<!-- - $\rho$: Spatial autocorrelation in residual errors (omitted variables) -->
<!-- - $\beta$: Regression coefficient of year on the variable -->
<!-- - se: standard error of $\beta$ -->
<!-- - t: test statistic -->
<!-- - p-value: p-value -->
<!-- - sig: whether p is < 0.05 -->
<!-- - reverse: whether the metric has been reversed so that lower numbers are better and higher numbers are worse (like food insecurity) -->
<!-- - outcome: if the regression is significant, is it getting better or worse? Accounts for the 'reverse' variable -->