---
title: "Coverage"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  purrr,
  broom,
  ggplot2,
  ggpubr,
  tibble,
  Hmisc,
  reshape2,
  viridisLite,
  conflicted,
  readr,
  sf,
  tmap,
  mapview,
  snakecase
)

# GitHub packages
pacman::p_load_current_gh(
  'ChrisDonovan307/projecter',
  'Food-Systems-Research-Institute/SMdata'
)

# Load SMdocs package
devtools::load_all()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

# Set a ggplot theme
dp_theme <- theme_classic() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.title.position = "top",
    panel.grid = element_blank(),
    text = element_text(family = "Times New Roman"),
    axis.title = element_text(family = "Times New Roman"),
    axis.text = element_text(family = "Times New Roman"),
    legend.title = element_text(family = "Times New Roman", hjust = 0.5),
    legend.text = element_text(family = "Times New Roman")
  )
usethis::use_data(dp_theme, overwrite = TRUE)

# Set theme palette
dp_palette <- c(
  "Economics"  = "#BF7417FF",
  "Environment"  = "#A8BE74FF",
  "Health" = "#9CB6A9FF",
  "Production" = "#635C72FF",
  "Social" = "#784116FF"
)
usethis::use_data(dp_palette, overwrite = TRUE)
```

Here we explore how secondary data sources cover the region by dimension, geography, and over time.


## Dimensional

Explore coverage by dimension. Things we are interested in:

- Proportion of indicators in each dimension that have at least one metric
- Same for county metric vs state
- Which indicators have no metrics

Indicators without any metrics:

```{r}
SMdocs::dp_meta %>% 
  filter(is.na(variable_name)) %>% 
  select(dimension, indicator) %>% 
  unique() %>% 
  get_reactable()
```

Proportion of indicator coverage by dimension:

```{r}
SMdocs::dp_meta %>% 
  group_by(dimension, indicator) %>% 
  summarize(
    has_metric = any(!is.na(variable_name)),
    .groups = 'drop_last'
  ) %>% 
  summarize(
    n_no_metric = sum(!has_metric),
    total_indicators = n(),
    prop_no_metric = mean(!has_metric),
    .groups = "drop"
  ) %>% 
  mutate(prop_no_metric = format(round(prop_no_metric, 3), nsmall = 3)) %>% 
  get_reactable()
```

How many metrics are state vs county:

```{r}
dat <- SMdocs::dp_meta %>% 
  filter(
    !is.na(variable_name),
    metric != 'NONE'
  ) %>% 
  select(
    dimension,
    metric,
    resolution
  ) %>% 
  mutate(resolution = case_when(
    resolution %in% c('state', 'system') ~ 'state',
    .default = 'county'
  ))

# Overall
overall_county <- mean(dat$resolution == 'county') * 100
overall_state <- 100 - overall_county
```

Overall, `{r} overall_county`% of metrics were at the county level, and `{r} overall_state`% were at the state level.

Which ones were state level:

```{r}
get_reactable(dat)
```

Break down by dimension:

```{r}
# By Dimension:
dat %>% 
  group_by(dimension) %>% 
  summarize(
    prop_county = mean(resolution == 'county'),
    prop_state = mean(resolution == 'state')
  ) %>% 
  mutate(across(starts_with('prop'), ~ format(round(.x, 3), nsmall = 3))) %>% 
  get_reactable()
```

## Geographic

### Summary Stats

Things we are interested in:

- How many metrics were represented in each county? Breakdown by dimension
- Which counties have least metrics
- Which metrics are only at state level

Start with metrics in each county, including population to poke around a bit with:

```{r}
#| label: geo_summary_stats
#| output: false
# get_str(SMdocs::dp_metrics_county)
n_county_metrics <- SMdocs::dp_metrics_county$variable_name %>% 
  unique %>% 
  length

# Latest population of each county
pops <- SMdocs::dp_weights %>% 
  filter(variable_name == 'population5YearSmooth', year == '2023') %>% 
  select(fips, value)

# Metrics per county
# get_str(SMdocs::dp_metrics_county)
tab_metrics <- SMdocs::dp_metrics_county %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name))) %>% 
  ungroup() %>% 
  left_join(SMdata::fips_key) %>% 
  select(-state_code) %>% 
  left_join(pops) %>% 
  rename(population_2023 = value) %>% 
  mutate(population_2023 = as.numeric(population_2023)) %>% 
  arrange(n_metrics)
get_reactable(tab_metrics)
```

Now break down by dimension:

```{r}
dims <- SMdocs::dp_meta %>% 
  select(variable_name, dimension) %>% 
  na.omit() %>% 
  filter(variable_name != 'NONE')
tab_metrics_dim <- SMdocs::dp_metrics_county %>% 
  left_join(dims) %>% 
  group_by(fips, dimension) %>% 
  summarize(n_metrics = length(unique(variable_name)))
  # distinct(fips) %>%
  # left_join(pops, by = 'fips') %>% 
  # rename(population_2023 = value)
get_reactable(tab_metrics_dim)
tab_metrics_dim
```


### Maps?

Turns out this map is not as interesting as I'd hoped it would be. Currently showing the unique count of metrics with any values in a county. Need a better way to represent the missing data here though. 

```{r}
#| output: false
# Load datasets from eda script
## Get bounding box - commenting out to avoid downloading this each time
# bbox_new <- st_bbox(neast_counties_2024)
# xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
# yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
# 
# bbox_new[1] <- bbox_new[1] - (0.05 * xrange) # xmin - left
# bbox_new[3] <- bbox_new[3] + (0.05 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.05 * yrange) # ymin - bottom
# bbox_new[4] <- bbox_new[4] + (0.05 * yrange) # ymax - top
# 
# tiles <- get_tiles(
#   bbox_new,
#   provider = "CartoDB.PositronNoLabels",
#   zoom = 7,
#   crop = TRUE
# )
# terra::saveRDS(tiles, 'data/data_paper/neast_tiles.rds')

# Load tiles for background
tiles <- terra::readRDS('data/data_paper/neast_tiles.rds')

# Counts of unique metrics in each county
# var_counts <- datasets$dp_metrics_all %>% 
var_counts <- dp_metrics_county %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name)))
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)

tmap_mode('plot')
map <- tm_shape(tiles) +
  tm_rgb() +
  tm_shape(df) +
  tm_polygons(
    "n_metrics", 
    palette = "brewer.greens",
    title = "Metric\nRepresentation",
    breaks = c(seq(21, 41, 5), 42:47),
    fill.legend = tm_legend(
      reverse = TRUE
    )
  ) +
  tm_borders(col = 'black', lwd = 1.25) +
  tm_layout(
    legend.position = c('left', 'top'),
    legend.title.fontface = 'bold',
    legend.width = 8,
    legend.height = 12,
    legend.title.size = 1.1,
    inner.margins = rep(0, 4),
    outer.margins = rep(0, 4),
    legend.text.size = 1
  )

# tmap_save(
#   tm = map,
#   filename = 'outputs/metric_coverage_map.png',
#   asp = 0,
#   dpi = 300
# )
```

```{r}
#| echo: false
#| fig-height: 8
#| fig-width: 8
#| out-width: '75%'
#| fig-align: 'center'
map
```

Let's try an interactive map so we can explore each dimension

```{r}
#| label: var_counts
#| output: false
# How many total metrics at county level
total_count <- dp_metrics_county %>% 
  pull(variable_name) %>% 
  unique %>% 
  length
total_count

# Dimension crosswalk
crosswalk <- dp_meta %>% 
  select(dimension, variable_name)

dimension_counts <- dp_meta %>% 
  select(dimension, variable_name, resolution) %>% 
  filter(!is.na(variable_name), resolution != 'state') %>% 
  select(-resolution) %>% 
  group_by(dimension) %>% 
  summarize(count = n())
dimension_counts
  
# Counts of unique metrics in each county by dimension
var_counts <- dp_metrics_county %>% 
  left_join(crosswalk) %>% 
  group_by(dimension, fips) %>% 
  summarize(
    n_metrics = length(unique(variable_name))
  ) %>% 
  left_join(dimension_counts) %>% 
  mutate(prop_metrics = n_metrics / count)
var_counts
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)
get_str(df)

maps <- df %>%
  split(.$dimension) %>%
  imap(~ {
    to_hide <- ifelse(.y == 'economics', FALSE, TRUE)
    mapview(
      .x, 
      layer.name = .y, 
      zcol = "prop_metrics",
      hide = to_hide,
      # col.regions = brewer.pal(5, "Greens"),
      col.regions = rev(viridis(5)),
      alpha.regions = 0.7
    )
  }) 
```

```{r}
#| echo: false
#| out-width: '75%'
Reduce(`+`, maps)
```

Hamilton, Warren counties missing data in NY, and Cameron county in PA stick out in particular. And lots of economic data missing from much of Maine and New Hampshire.

## Temporal

### Indicators

Total number of indicators represented each year.

```{r}
#| label: Total indicators each year
#| out-width: 75%
#| fig-align: center
# Join based on 'variable_name'
merged_df <- SMdocs::dp_metrics_county %>%
  left_join(SMdocs::dp_meta %>%
              select(variable_name,
                     dimension = dimension,
                     index = index,
                     indicator = indicator
              ),
            by = "variable_name") 

# Filter out before 2000 and NAs; create "indicator count"
indicators_by_year_data <- merged_df %>%
  filter(!is.na(indicator)) %>%
  filter(year >= 2000) %>%
  distinct(indicator, dimension, year) %>%
  count(year, dimension, name = "indicator_count") %>% 
  mutate(year = as.numeric(year))
# get_str(indicators_by_year_data)

# Histogram
indicator_plot <- ggplot(
    indicators_by_year_data, 
    aes(x = year, y = indicator_count, fill = dimension)
  ) +
  geom_col(position = "stack") +
  scale_y_continuous(breaks = seq(0, max(1000), by = 5)) + 
  # max is set to 1000 values - change max for more data
  scale_x_continuous(
    breaks = seq(
      min(indicators_by_year_data$year), 
      max(indicators_by_year_data$year),
      by = 5)) +
  scale_fill_manual(
    values = dp_palette,
    labels = to_title_case
  ) +
  labs(
    x = "Year",
    y = "Number of Indicators Represented",
    fill = "Dimension"
  ) +
  dp_theme

# Save to outputs
ggsave(
  indicator_plot,
  filename = 'outputs/fig_indicator_coverage.png',
  width = 5,
  height = 4,
  units = 'in',
  dpi = 300,
  bg = 'white'
)

# Render to Quarto
indicator_plot
```

### Metrics

Metric representation each year.

```{r}
#| label: Metric representation each year
#| fig-height: 8
#| fig-width: 7
# Join based on "variable_name"
metrics_and_meta <- SMdocs::dp_metrics %>%
  left_join(SMdocs::dp_meta %>%
              select(variable_name,
                     dimension,
                     index,
                     indicator,
                     metric),
            by = "variable_name") 

# Group dimension columns together in ascending order; filter out before 2000 
# and NAs
heatmap_easy_data <- metrics_and_meta %>%
  filter(!is.na(indicator)) %>%
  filter(year >= 2000) %>%
  distinct(indicator, dimension, year, metric)

heatmap_easy_data <- heatmap_easy_data %>%
  mutate(
    dimension = factor(dimension, levels = sort(unique(dimension)))
  ) %>%
  filter(!is.na(metric)) %>%
  arrange(dimension, metric) %>%
  mutate(
    metric = factor(metric, levels = rev(unique(metric)))
  )

heatmap_easy_data <- heatmap_easy_data %>%
  distinct(year, dimension, metric)

# Full grid of all combinations (to show blanks)
full_grid <- expand.grid(
  metric = unique(heatmap_easy_data$metric),
  year = unique(heatmap_easy_data$year),
  stringsAsFactors = FALSE
)

# Left join to get "present" boxes
data_full <- left_join(full_grid, heatmap_easy_data, by = c("metric", "year"))

# Plot
metric_heatmap <- ggplot(
  data_full, 
  aes(x = as.factor(year), y = metric, fill = dimension)
) +
  geom_tile(color = "gray80", linewidth = 0.25) +
  scale_y_discrete(labels = to_title_case) +
  scale_x_discrete(
    breaks = seq(
      min(heatmap_easy_data$year), 
      max(heatmap_easy_data$year), 
      by = 5)
  ) +
  scale_fill_manual(
    values = c(dp_palette, "FALSE" = "white"),
    na.value = "white",
    labels = to_title_case
  ) +
  labs(
    x = "Year",
    y = "Metric",
    fill = "Dimension"
  ) +
  dp_theme

# Save to outputs for paper
ggsave(
  metric_heatmap,
  filename = 'outputs/fig_metric_heatmap.png',
  width = 7,
  height = 8,
  units = 'in',
  dpi = 300,
  bg = 'white'
)

# Render to Quarto
metric_heatmap
```
