---
title: "Coverage"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  purrr,
  broom,
  ggplot2,
  ggpubr,
  tibble,
  Hmisc,
  reshape2,
  viridisLite,
  conflicted,
  readr,
  sf,
  tmap,
  mapview
)

# GitHub packages
pacman::p_load_gh(
  'ChrisDonovan307/projecter',
  'Food-Systems-Research-Institute/SMdata'
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')
source('dev/data_pipeline_functions.R')

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
datasets <- readRDS('data/data_paper/datasets_from_eda.rds')
```

Here we explore how secondary data sources cover the region by dimension, geography, and over time.


## Dimensional

Explore coverage by dimension.

## Geographic

Turns out this map is not as interesting as I'd hoped it would be. Currently showing the unique count of metrics with any values in a county. Need a better way to represent the missing data here though. 

```{r}
#| output: false
# Load datasets from eda script
## Get bounding box - commenting out to avoid downloading this each time
# bbox_new <- st_bbox(neast_counties_2024)
# xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
# yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
# 
# bbox_new[1] <- bbox_new[1] - (0.05 * xrange) # xmin - left
# bbox_new[3] <- bbox_new[3] + (0.05 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.05 * yrange) # ymin - bottom
# bbox_new[4] <- bbox_new[4] + (0.05 * yrange) # ymax - top
# 
# tiles <- get_tiles(
#   bbox_new,
#   provider = "CartoDB.PositronNoLabels",
#   zoom = 7,
#   crop = TRUE
# )
# terra::saveRDS(tiles, 'data/data_paper/neast_tiles.rds')

# Load tiles for background
tiles <- terra::readRDS('data/data_paper/neast_tiles.rds')

# Counts of unique metrics in each county
var_counts <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name)))
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)

tmap_mode('plot')
map <- tm_shape(tiles) +
  tm_rgb() +
  tm_shape(df) +
  tm_polygons(
    "n_metrics", 
    palette = "brewer.greens",
    title = "Metric\nRepresentation",
    breaks = seq(21, 45, 3),
    fill.legend = tm_legend(
      reverse = TRUE
    )
  ) +
  tm_borders(col = 'black', lwd = 1.25) +
  tm_layout(
    legend.position = c('left', 'top'),
    legend.title.fontface = 'bold',
    legend.width = 8,
    legend.height = 12,
    legend.title.size = 1.1,
    inner.margins = rep(0, 4),
    outer.margins = rep(0, 4),
    legend.text.size = 1
  )

tmap_save(
  tm = map,
  filename = 'outputs/metric_coverage_map.png',
  asp = 0,
  dpi = 300
)
```

```{r}
#| echo: false
#| fig-height: 8
#| fig-width: 8
#| out-width: '75%'
#| fig-align: 'center'
map
```

Let's try an interactive map so we can explore each dimension

```{r}
#| label: var_counts
#| output: false
# How many total metrics at county level
total_count <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  pull(variable_name) %>% 
  unique %>% 
  length
total_count

# Dimension crosswalk
crosswalk <- data_paper_meta %>% 
  select(dimension, variable_name)

dimension_counts <- data_paper_meta %>% 
  select(dimension, variable_name, resolution) %>% 
  filter(!is.na(variable_name), resolution != 'state') %>% 
  select(-resolution) %>% 
  group_by(dimension) %>% 
  summarize(count = n())
dimension_counts
  
# Counts of unique metrics in each county by dimension
var_counts <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  left_join(crosswalk) %>% 
  group_by(dimension, fips) %>% 
  summarize(
    n_metrics = length(unique(variable_name))
  ) %>% 
  left_join(dimension_counts) %>% 
  mutate(prop_metrics = n_metrics / count)
var_counts
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)
get_str(df)

maps <- df %>%
  split(.$dimension) %>%
  imap(~ {
    to_hide <- ifelse(.y == 'economics', FALSE, TRUE)
    mapview(
      .x, 
      layer.name = .y, 
      zcol = "prop_metrics",
      hide = to_hide,
      # col.regions = brewer.pal(5, "Greens"),
      col.regions = rev(viridis(5)),
      alpha.regions = 0.7
    )
  }) 
```

```{r}
#| echo: false
#| out-width: '75%'
Reduce(`+`, maps)
```

Hamilton, Warren counties missing data in NY, and Cameron county in PA stick out in particular. And lots of economic data missing from much of Maine and New Hampshire.

## Temporal

Add Isabella's graphs showing metrics and indicators over the years.

