---
title: "Exploratory Data Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  knitr,
  skimr,
  ggplot2,
  ggpubr,
  psych,
  tibble,
  tmap,
  maptiles,
  tmaptools,
  sf,
  Hmisc,
  reshape2,
  mapview,
  viridisLite
)

pacman::p_load_gh(
  'ChrisDonovan307/projecter',
  'Food-Systems-Research-Institute/SMdata'
)

# Functions
source('dev/get_reactable.R')
source('dev/get_missing.R')
source('dev/data_pipeline_functions.R')

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```


## Wrangling

```{r}
#| output: false
#| freeze: true
# Create a list of the clean datasets we will use throughout analysis
datasets <- list()

# Check out summary from OneDrive that has our variable names
get_str(data_paper_meta)

# Pull vars to reduce metrics down
meta <- data_paper_meta %>% 
  filter(
    !is.na(variable_name),
    !variable_name %in% c('tbd', '', 'NONE'),
    analyze != '',
    !is.na(analyze)
  ) %>% 
  select(
    variable_name,
    resolution,
    starts_with('n_'),
    latest_year,
    dimension:metric,
    weighting:desirable,
  )
get_str(meta)

# Load neast metrics for data paper, both county and state
get_str(data_paper_metrics)
all <- data_paper_metrics %>% 
  filter_fips('neast') %>% 
  inner_join(select(meta, variable_name, resolution)) %>% 
  mutate(across(c(year, value), as.numeric))
get_str(all)

## Cleaning
# "annualAvgWklyWageNAICS11" should have NAs instead of zeroes
all$value[all$variable_name == "annualAvgWklyWageNAICS11" & all$value == 0] <- NA

# Save to dataset list
datasets$dp_metrics_all <- all

# Pull out county and state separately, based on spec in metadata
dp_county <- all %>% 
  filter_fips('counties') %>% 
  filter(resolution != 'state') %>% 
  select(-resolution)
get_str(dp_county)
unique(dp_county$fips)
datasets$dp_metrics_county <- dp_county

dp_state <- all %>% 
  filter_fips('states') %>% 
  filter(resolution == 'state') %>% 
  select(-resolution)
get_str(dp_state)
unique(dp_state$fips)
datasets$dp_metrics_state <- dp_state

# Get another DF of weighting variables
weights <- metrics %>%
  filter(variable_name %in% weighting$variable_name) %>%
  filter_fips('neast')
get_str(weights)
 
# Save to dataset list
datasets$dp_weights <- weights
  
# Pivot wider for transformations, then alphabetize columns
dp_county_wide <- dp_county %>% 
  # Remove one weird doubled up value
  filter(!(
    fips == '42' & year == 2022 & variable_name == 'hayYieldMeasuredInTonsAcre'
  )) %>%
  pivot_wider(
    id_cols = c(fips, year),
    values_from = 'value',
    names_from = 'variable_name'
  ) %>% 
  select(fips, year, sort(names(.)[2:length(names(.))]))
get_str(dp_county_wide)

# Save to datasets
datasets$dp_county_wide <- dp_county_wide
```


## Missing Data

Explore missing data for each individual metric. Note that there is a baseline of missing for every one because of the issues with Connecticut counties. This is a bit awkward.

```{r}
#| label: missing_data
#| output: false
skimr::skim(dp_county_wide)
# Missing years are county areas, that's fine
# Other missing values are because years don't match up, that's fine
get_str(dp_county_wide)
vars <- names(dp_county_wide[, 3:ncol(dp_county_wide)])

out <- map_dbl(vars, ~ {
  years <- dp_county_wide %>% 
    drop_na(.x) %>% 
    distinct(year) %>% 
    pull(year)
  df <- dp_county_wide %>% 
    select(year, fips, all_of(.x)) %>% 
    filter(year %in% years) %>% 
    complete(year, fips)
  prop_mis <- sum(is.na(df[.x])) / nrow(df)
  return(prop_mis)
})

# Graph this
miss <- bind_cols(vars, out) %>% 
  setNames(c('var', 'prop_miss'))
miss_plot <- miss %>% 
  ggplot(aes(y = reorder(var, prop_miss), x = prop_miss)) +
  geom_col(
    color = 'black',
    fill = '#154734'
  ) +
  labs(
    x = 'Proportion Missing',
    y = 'Metric'
  ) +
  theme_bw()
```

```{r}
#| label: plot_missing
#| echo: false
#| fig-height: 8
#| fig-width: 6
miss_plot
```


## Distributions

Exploring distributions using latest year available for each metric.

### County Data

```{r}
#| label: metric_distributions_county
#| fig-height: 25
#| fig-width: 10
#| freeze: true
# Get skews of variables
dat <- dp_county %>% 
  get_latest_year() %>% 
  pivot_wider(
    id_cols = fips,
    values_from = 'value',
    names_from = 'variable_name'
  )
# get_str(dat)

skewed <- psych::describe(dat[, -1]) %>% 
  as.data.frame() %>% 
  rownames_to_column('variable_name') %>% 
  dplyr::select(variable_name, skew) %>% 
  dplyr::filter(abs(skew) > 2) %>% 
  pull(variable_name)

plots <- map(names(dat)[-1], \(var){
  # color based on skewness
  if (var %in% skewed) {
    fill <- 'red'
    color <- 'darkred'
  } else {
    fill <- '#154724'
    color <- 'black'
  }
  
  # Make plot for variable
  dat %>% 
    ggplot(aes(x = !!sym(var))) + 
    geom_density(
      fill = fill,
      color = color,
      alpha = 0.5
    ) +
    theme_classic() +
    theme(plot.margin = unit(c(rep(0.5, 4)), 'cm'))
}) 

# Arrange them in 4 columns
ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```


### State Data

```{r}
#| label: state_distributions
#| fig-height: 10
#| fig-width: 10
source('dev/get_distributions.R')
dat <- dp_state %>% 
  filter(!(
    fips == '42' & year == 2022 & variable_name == 'hayYieldMeasuredInTonsAcre'
  )) %>% 
  get_latest_year() %>% 
  pivot_wider(
    id_cols = fips,
    values_from = 'value',
    names_from = 'variable_name'
  )
get_distributions(
  df = dat,
  n_col = 4,
  n_row = 5
)
```


## Map of Coverage

Turns out this map is not as interesting as I'd hoped it would be. Currently showing the unique count of metrics with any values in a county. Need a better way to represent the missing data here though. 

```{r}
#| output: false
## Get bounding box - commenting out to avoid downloading this each time
# bbox_new <- st_bbox(neast_counties_2024)
# xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
# yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
# 
# bbox_new[1] <- bbox_new[1] - (0.05 * xrange) # xmin - left
# bbox_new[3] <- bbox_new[3] + (0.05 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.05 * yrange) # ymin - bottom
# bbox_new[4] <- bbox_new[4] + (0.05 * yrange) # ymax - top
# 
# tiles <- get_tiles(
#   bbox_new,
#   provider = "CartoDB.PositronNoLabels",
#   zoom = 7,
#   crop = TRUE
# )
# terra::saveRDS(tiles, 'data/data_paper/neast_tiles.rds')

# Load tiles for background
tiles <- terra::readRDS('data/data_paper/neast_tiles.rds')

# Counts of unique metrics in each county
var_counts <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name)))
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)

tmap_mode('plot')
map <- tm_shape(tiles) +
  tm_rgb() +
  tm_shape(df) +
  tm_polygons(
    "n_metrics", 
    palette = "brewer.greens",
    title = "Metric\nRepresentation",
    breaks = seq(21, 45, 3),
    fill.legend = tm_legend(
      reverse = TRUE
    )
  ) +
  tm_borders(col = 'black', lwd = 1.25) +
  tm_layout(
    legend.position = c('left', 'top'),
    legend.title.fontface = 'bold',
    legend.width = 8,
    legend.height = 12,
    legend.title.size = 1.1,
    inner.margins = rep(0, 4),
    outer.margins = rep(0, 4),
    legend.text.size = 1
  )

tmap_save(
  tm = map,
  filename = 'outputs/metric_coverage_map.png',
  asp = 0,
  dpi = 300
)
```

```{r}
#| echo: false
#| fig-height: 8
#| fig-width: 8
map
```

Let's try an interactive map so we can explore each dimension

```{r}
#| label: var_counts
#| output: false
# How many total metrics at county level
total_count <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  pull(variable_name) %>% 
  unique %>% 
  length
total_count

# Dimension crosswalk
crosswalk <- data_paper_meta %>% 
  select(dimension, variable_name)

dimension_counts <- data_paper_meta %>% 
  select(dimension, variable_name, resolution) %>% 
  filter(!is.na(variable_name), resolution != 'state') %>% 
  select(-resolution) %>% 
  group_by(dimension) %>% 
  summarize(count = n())
dimension_counts
  
# Counts of unique metrics in each county by dimension
var_counts <- datasets$dp_metrics_all %>% 
  filter_fips('counties') %>% 
  left_join(crosswalk) %>% 
  group_by(dimension, fips) %>% 
  summarize(
    n_metrics = length(unique(variable_name))
  ) %>% 
  left_join(dimension_counts) %>% 
  mutate(prop_metrics = n_metrics / count)
var_counts
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)
get_str(df)

maps <- df %>%
  split(.$dimension) %>%
  imap(~ {
    to_hide <- ifelse(.y == 'economics', FALSE, TRUE)
    mapview(
      .x, 
      layer.name = .y, 
      zcol = "prop_metrics",
      hide = to_hide,
      # col.regions = brewer.pal(5, "Greens"),
      col.regions = rev(viridis(5)),
      alpha.regions = 0.7
    )
  }) 
```

```{r}
#| echo: false
Reduce(`+`, maps)
```

Hamilton, Warren counties missing data in NY, and Cameron county in PA stick out in particular. And lots of economic data missing from much of Maine and New Hampshire.


## Time Series

Plot time series of average variable that has at least 2 data points. These are raw, unweighted values.


### County Data

```{r}
#| label: prep_series
#| output: false
# Prepare vector of county variables that are time series
get_str(dp_county)
vars <- unique(dp_county$variable_name)
series_vars <- map(vars, ~ {
  count <- dp_county %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
datasets$series_vars <- series_vars
```

A few hinky things going on here - voter turnout is apparently flat across all counties. It looks like this is because it is only 2023 to 2025 and it is from a secondary source that pulled it from the 5-year census. So this needs to be fixed.

Land value per acre also appears jenky, but it is because Richmond County, NY (Staten Island) has a land value of 2.5 million per acre while most other counties are below 100k. We should probably remove this as an outlier.

```{r}
#| label: county_time_series
#| fig-height: 25
#| fig-width: 12
plots <- map(series_vars, ~ {
  df <- dp_county %>%
    filter(variable_name == .x) %>%
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
      # col = '#154734'
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```

```{r}
#| label: checking
#| eval: false
#| include: false
dp_county %>% 
  filter(variable_name == 'landValPerAcre', year == 2022) %>% 
  pull(value) %>% 
  hist()

dp_county %>% 
  filter(variable_name == 'landValPerAcre', year == 2022) %>% 
  arrange(value)
# 36085 has 2.5 million. Everything else belo2 600k, mostly below 100k

fips_key$county_name[fips_key$fips == 36085]
# Richmond County, NY
```


### State Data

```{r}
#| label: state_time_series
#| fig-height: 12
#| fig-width: 12
# get_str(dp_state)
vars <- unique(dp_state$variable_name)
state_series_vars <- map(vars, ~ {
  count <- dp_state %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()

# Save vector of series vars for later
datasets$state_series_vars <- state_series_vars

plots <- map(state_series_vars, ~ {
  df <- dp_state %>%
    filter(variable_name == .x) %>%
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
      # col = '#154734'
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 3,
  nrow = 4
)
```

```{r}
#| label: saveRDS
#| echo: false
# Save dataset list to use elsewhere
saveRDS(datasets, 'data/data_paper/datasets.rds')
```


## Correlations

```{r}
#| label: correlations
#| output: false
#| fig-height: 10
#| fig-width: 10

# metric crosswalk
crosswalk <- data_paper_meta %>% 
  select(variable_name, metric)

# Get df ready
get_str(metrics)
df <- metrics %>% 
  filter(variable_name %in% data_paper_meta$variable_name) %>% 
  get_latest_year() %>% 
  filter_fips('new') %>% 
  mutate(variable_name = str_split_i(variable_name, '_', 1)) %>% 
  left_join(crosswalk) %>% 
  select(-variable_name) %>% 
  pivot_wider(
    id_cols = fips,
    values_from = value,
    names_from = metric
  )
get_str(df)

# Get an ordered framework from which we will pull metrics in alphabetical order
# and get the counts of metrics per dimension that we will use to draw lines
# between dimensions
get_str(data_paper_meta)
ordered_framework <- data_paper_meta %>% 
  arrange(dimension, metric) %>% 
  filter(metric %in% names(df))
ordered_framework

# Get line placements to divide dimensions
# Reverse them to match alphabetical order of dimensions top to bottom
# Get cumulative sums to space them out across graph
# Then add 0.5 to each to put them in between cells
line_placements <- ordered_framework %>% 
  group_by(dimension) %>% 
  summarize(count = n()) %>% 
  pull(count) %>% 
  rev() %>% 
  cumsum() %>% 
  {. + 0.5}
line_placements

# Reorder our df in proper metric order
metric_order <- ordered_framework %>% 
  pull(metric)
df <- df %>% 
  select(fips, any_of(metric_order)) %>% 
  select(rev(everything()))
get_str(df)

# Create matrix of values
mat <- df %>% 
  na.omit() %>% 
  select(-fips) %>% 
  as.matrix()

# Get correlations
cor <- rcorr(mat, type = 'pearson')

# Melt correlation values and rename columns
cor_r <- reshape::melt(cor$r) %>% 
  setNames(c('var_1', 'var_2', 'value'))

# Save p values
cor_p <- melt(cor$P)
p.value <- cor_p$value

# Make heatmap
corplot <- cor_r %>% 
  ggplot(aes(var_1, var_2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(
    low = "#762a83", 
    mid = "white", 
    high = "#1b7837", 
    midpoint = 0
  ) +
  geom_hline(yintercept = line_placements[1]) +
  geom_hline(yintercept = line_placements[2]) +
  geom_hline(yintercept = line_placements[3]) +
  geom_hline(yintercept = line_placements[4]) +
  geom_vline(xintercept = line_placements[1]) +
  geom_vline(xintercept = line_placements[2]) +
  geom_vline(xintercept = line_placements[3]) +
  geom_vline(xintercept = line_placements[4]) +
  theme(
    axis.text.x = element_text(
      hjust = 1, 
      angle = 45
    )
    # axis.text.x = element_blank(),
    # axis.ticks.x = element_blank()
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = NULL
  )
corplot
ggsave(
  plot = corplot,
  filename = 'outputs/corr_plot.png',
  height = 8,
  width = 9.5,
  dpi = 300
)
```

```{r}
#| echo: false
#| fig-height: 8
#| fig-width: 10
corplot
```
