---
title: "Exploratory Data Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  purrr,
  broom,
  skimr,
  ggplot2,
  ggpubr,
  tibble,
  Hmisc,
  reshape2,
  viridisLite,
  conflicted,
  readr,
  reactable
)

# GitHub packages
pacman::p_load_current_gh(
  'ChrisDonovan307/projecter',
  'Food-Systems-Research-Institute/SMdata'
)

# Load SMdocs
devtools::load_all()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```


## New Missing Data

New plan for showing missing data in a more granular way:

- By metric: for each metric, figure out how many counties/states and years it should represent. Multiply. This gives total number of values it should include. Then compare that total number of data points.
- By state/county: figure out how many metrics should be in that county, and then for how many years (not including when we just don't have those years represented, which is different I suppose).

### By Metric

County metrics only

```{r}
#| label: mis_by_metric

# get_str(dp_metrics_county_wide)
# get_str(dp_metrics_county)
vars <- meta_vars(dp_metrics_county)

# Years covered by each metric, multiplied by number of counties: 209 (not CT)
n_counties <- 209

# Get years, number of data points, total possible number of points based on 
# years and 209 counties, then the percentage missing
metric_cov <- map(vars, ~ {
  var_df <- dp_metrics_county %>%
    filter(
      variable_name == .x,
      str_detect(fips, '^09', negate = TRUE)
    )
  n_years <- var_df %>% 
    pull(year) %>%
    unique() %>%
    length()
  n_points <- nrow(var_df)
  df <- data.frame(
    variable_name = .x,
    n_years = n_years, 
    n_points = n_points
  )
  return(df)
}) %>% 
  bind_rows() %>% 
  mutate(
    n_possible = n_years * n_counties,
    perc_miss = round(100 - ((n_points / n_possible) * 100), 2)
  )

missy_metric_df <- metric_cov %>% 
  filter(perc_miss > 5)
missy_metrics <- missy_metric_df$variable_name
print_missy_metrics <- paste0(missy_metrics, collapse = ', ')
  
get_reactable(metric_cov)
```

The metrics with > 5% missing are: `{r} print_missy_metrics`. Now see where they are missing data:

```{r}
#| label: missy_metrics_table
non_ct_fips <- fips_key %>% 
  filter(
    str_length(fips) == 5,
    str_detect(fips, '^09', negate = TRUE)
  ) %>% 
  pull(fips)

missing_table <- dp_metrics_county %>% 
  filter(
    variable_name %in% missy_metrics,
    str_detect(fips, '^09', negate = TRUE)
  ) %>% 
  group_by(variable_name, year) %>%
  summarize(
    fips_present = c(fips),
    fips_missing = list(setdiff(non_ct_fips, fips_present)),
  ) %>% 
  select(-fips_present) %>%
  unique() %>% 
  mutate(
    n_fips_missing = lengths(fips_missing)
  )

# Table
get_reactable(
  missing_table,
  defaultColDef = reactable::colDef(minWidth = 100),
  columns = list(
    fips_missing = reactable::colDef(minWidth = 500)
  )
)
```

### By County

What are the most missed counties? For each metric, rank most missing counties in order of times they are missing

```{r}
#| label: missed_counties
out <- map(missy_metrics, ~ {
  missing_table %>% 
    filter(variable_name == .x) %>% 
    pull(fips_missing) %>% 
    unlist() %>% 
    table() %>% 
    as.data.frame() %>% 
    setNames(c('fips', 'times_missing')) %>% 
    arrange(desc(times_missing))
}) %>% 
  setNames(c(missy_metrics))

# Add county names
out <- map(out, ~ {
  left_join(.x, select(fips_key, -state_code))
})
```

Conservation Easement Acres:

```{r}
#| class: centered-table
get_reactable(out$consEasementAcres, fullWidth = FALSE)
```

Proportion of Invasive Species:

```{r}
#| class: centered-table
get_reactable(out$propInvasive)
```

Proportion of farms with precision ag:

```{r}
#| class: centered-table
get_reactable(out$propOpsPrecision)
```

Tree Size Diversity:

```{r}
#| class: centered-table
get_reactable(out$sizeDiversity)
```

Tree Species Diversity:

```{r}
#| class: centered-table
get_reactable(out$treeDiversity)
```


## Old Missing Data

Explore missing data for each individual metric. Note that there is a baseline of missing for every one because of the issues with Connecticut counties. This is a bit awkward.

```{r}
#| label: missing_data
#| output: false
skimr::skim(SMdocs::dp_metrics_county_wide)
# Missing years are county areas, that's fine
# Other missing values are because years don't match up, that's fine
get_str(dp_metrics_county_wide)
vars <- names(dp_metrics_county_wide[, 3:ncol(dp_metrics_county_wide)])

out <- map_dbl(vars, ~ {
  years <- dp_metrics_county_wide %>% 
    drop_na(.x) %>% 
    distinct(year) %>% 
    pull(year)
  df <- dp_metrics_county_wide %>% 
    select(year, fips, all_of(.x)) %>% 
    filter(year %in% years) %>% 
    complete(year, fips)
  prop_mis <- sum(is.na(df[.x])) / nrow(df)
  return(prop_mis)
})

# Graph this
miss <- bind_cols(vars, out) %>% 
  setNames(c('var', 'prop_miss'))
miss_plot <- miss %>% 
  ggplot(aes(y = reorder(var, prop_miss), x = prop_miss)) +
  geom_col(
    color = 'black',
    fill = '#154734'
  ) +
  labs(
    x = 'Proportion Missing',
    y = 'Metric'
  ) +
  theme_bw()
```

```{r}
#| label: plot_missing
#| echo: false
#| out-width: 75%
#| fig-align: 'center'
miss_plot
```

## Distributions

Exploring distributions using latest year available for each metric.

### County Data

```{r}
#| label: metric_distributions_county
#| fig-height: 25
#| fig-width: 10
#| cache: true
# Get skews of variables
# get_str(dp_metrics_county)
dat <- dp_metrics_county %>% 
  get_latest_year() %>% 
  pivot_wider(
    id_cols = fips,
    values_from = 'value',
    names_from = 'variable_name'
  )
# get_str(dat)

skewed <- psych::describe(dat[, -1]) %>% 
  as.data.frame() %>% 
  mutate(across(everything(), as.numeric)) %>% 
  tibble::rownames_to_column('variable_name') %>% 
  dplyr::select(variable_name, skew) %>% 
  dplyr::filter(abs(skew) > 1) %>% 
  dplyr::pull(variable_name)

plots <- map(names(dat)[-1], \(var){
  # color based on skewness
  if (var %in% skewed) {
    fill <- 'red'
    color <- 'darkred'
  } else {
    fill <- '#154724'
    color <- 'black'
  }
  
  # Make plot for variable
   dat %>% 
    select(!!sym(var)) %>% 
    mutate(!!sym(var) := as.numeric(!!sym(var))) %>% 
    na.omit() %>% 
    ggplot(aes(x = !!sym(var), group = 1)) + 
    geom_density(
      fill = fill,
      color = color,
      alpha = 0.5
    ) +
    theme_classic() +
    scale_x_continuous(n.breaks = 5) +
    theme(plot.margin = unit(c(rep(0.5, 4)), 'cm'))
}) 

# Arrange them in 4 columns
# TODO: Turn this into a package function
n_col <- 4
n_plots <- length(plots)
n_rows <- ceiling(n_plots/n_col)
ggarrange(
  plotlist = plots,
  ncol = n_col,
  nrow = n_rows
)
```


### State Data

```{r}
#| label: state_distributions
#| fig-height: 10
#| fig-width: 10
#| cache: true
state_dat <- dp_metrics_state %>% 
  filter(!(
    fips == '42' & year == 2022 & variable_name == 'hayYieldMeasuredInTonsAcre'
  )) %>% 
  get_latest_year() %>% 
  pivot_wider(
    id_cols = fips,
    values_from = 'value',
    names_from = 'variable_name'
  )
state_plots <- SMdocs::get_distributions(df = state_dat)
  
# Arrange with proper number of rows based on given number of cols
n_col <- 4
n_plots <- length(state_plots)
n_rows <- ceiling(n_plots/n_col)
ggpubr::ggarrange(
  plotlist = state_plots,
  ncol = n_col,
  nrow = n_rows
)
```
