---
title: "Regional Analyses"
format:
  html:
    fig-dpi: 200
editor_options: 
  chunk_output_type: console
warnings: false
engine: knitr
execute:
  cache: false
filters:
    - ../dev/show-lang.lua
---

::: {.callout-note collapse='false' title='Caution'}
The Sustainability Metrics project, as well as this site itself, are works in progress. All data and analyses shown here are preliminary. 
:::

```{r}
#| label: setup
#| echo: false
#| cache: false
pacman::p_load(
  dplyr,
  conflicted,
  purrr,
  stringr,
  ggraph,
  igraph,
  dplyr,
  readr,
  RColorBrewer,
  paletteer,
  snakecase,
  tidyr,
  htmltools,
  reactable,
  downloadthis,
  reticulate
)

pacman::p_load_gh(
  'ChrisDonovan307/projecter',
  'Food-Systems-Research-Institute/SMdata'
)

# Functions
source('dev/get_dimension_ggraph.R')
source('dev/data_pipeline_functions.R')

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::arrange(),
  dplyr::summarize(),
  dplyr::as_data_frame(),
  .quiet = TRUE
)
```


## Introduction {#sec-intro}

This group of pages will present a survey of secondary data at the regional/county level in the Northeast. Note that the graphs and CSV below have been updated to reflect some liberties and consolidation we have done with the framework to make it a bit more presentable. 

## Dimension Graphs

```{r}
#| label: prep_graphs
#| include: false
# Filter DF and check data
source('dev/get_reactable.R')
tree <- data_paper_tree %>% 
  mutate(
    indicator = case_when(
        str_detect(indicator, 'inputs bought locally') ~ '% local inputs',
        str_detect(indicator, 'production inputs') ~ 'production inputs',
        str_detect(indicator, 'metabolic') ~ 'presence of disease',
        .default = indicator
      ) %>% 
      str_remove(' of food business')
  )
# get_reactable(tree)

# Set up list of results
plots <- list()
```

### Economics

```{r}
#| fig-height: 5
#| fig-width: 12
plots$econ <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'economics',
  y_limits = c(-3.75, 3.25),
  palette = "scico::batlowW"
)
plots$econ
```

### Environment

```{r}
#| fig-height: 7
#| fig-width: 10
plots$env <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'environment',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$env
```

### Health

```{r}
#| fig-height: 7
#| fig-width: 12
plots$health <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'health',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$health
```

### Production

```{r}
#| fig-height: 5
#| fig-width: 14
plots$prod <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'production',
  y_limits = c(-4.5, 3.25),
  palette = "scico::batlowW"
)
plots$prod
```

### Social

```{r}
#| fig-height: 8
#| fig-width: 13
plots$social <- get_dimension_ggraph(
  framework_df = tree,
  include_metrics = TRUE,
  dimension_in = 'social',
  y_limits = c(-3, 3.25),
  palette = "scico::batlowW"
)
plots$social
```

```{r}
#| label: save_pngs
#| include: false
# NOT SHOWN - using this to save some PNGs of graphs
# df of dimensions to use
dims <- data.frame(
  heights = c(5, 7, 6, 5, 6),
  widths = c(15, 10, 10, 12, 13)
)
pwalk(list(plots, names(plots), 1:nrow(dims)), \(plot, name, i) {
  ggsave(
    paste0('outputs/dimension_', name, 'graph.png'),
    plot = plot,
    height = dims[i, 1],
    width = dims[i, 2],
    units = 'in',
    bg = 'white',
    dpi = 500
  )  
})
```


## Sunburst

Little bit messy but maybe fun. Need to fix labels.

```{python}
import plotly.io as pio
import plotly.express as px
import plotly.graph_objects as go

# Set renderer to notebook
pio.renderers.default = "notebook_connected"

# Pull data from R
tree = r.tree

# Create plot
px.sunburst(
  tree,
  path=["dimension", "index", "indicator", "metric"]
)
```


## Dimension Table

Easier to interpret table with the framework.

```{r}
data_paper_tree %>% 
  mutate(metric = case_when(
    str_detect(metric, 'NONE') ~ NA,
    .default = metric
  )) %>% 
  arrange(dimension, index, indicator, metric) %>% 
  get_reactable()
```

