---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: true
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  conflicted,
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  knitr,
  kableExtra,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  spatialreg,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr,
  lmtest,
  sandwich,
  performance,
  snakecase,
  ggtext
)

pacman::p_load_current_gh(
  'ChrisDonovan307/projecter', 
  'Food-Systems-Research-Institute/SMdata'
)

# Load SMdocs project
devtools::load_all()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```

Prep graphing functions:

```{r}
#| label: graph_functions
## Functions to make regression CI graphs

# Take DF of model outputs (variable_name, coefficient, p values, outcomes...)
# and format for regression graphs. Returns formatted data as well as a lookup
# to match "indicator-metric" label to dimension to add colors.
prepare_figure_data <- function(model_output_df) {
  # Title case and combine labels
  figure_data <- model_output_df %>%
    mutate(
      indicator = snakecase::to_title_case(indicator),
      metric = snakecase::to_title_case(metric),
      combined_label = paste(indicator, "-", metric)
    )
  
  # Order dimensions alphabetically
  ordered_dims <- sort(unique(figure_data$dimension))
  
  # Arrange and filter
  figure_data <- figure_data %>%
    mutate(
      dimension = factor(dimension, levels = ordered_dims)
    ) %>%
    arrange(dimension, estimate) %>%
    mutate(
      combined_label = factor(combined_label, levels = rev(unique(combined_label)))
    ) %>%
    # filter(metric != "Poor Mental Health Days") %>%
    ungroup()
  
  # Create label-dimension lookup
  label_dimension_lookup <- figure_data %>%
    distinct(combined_label, dimension)
  
  # Return both processed data and label-dimension
  list(
    figure_data = figure_data,
    label_dimension_lookup = label_dimension_lookup
  )
}

# Create a function for making y-axis labels using a palette for text colors
# and label dimension lookup from prepare_figure_data()
make_y_axis_labels <- function(label_dimension_lookup, dp_text_palette) {
  function(x) {
    sapply(x, function(y) {
      dim_val <- label_dimension_lookup$dimension[label_dimension_lookup$combined_label == y]
      dimension_label_color <- dp_text_palette[as.character(dim_val)]
      
      # Split the combined label into indicator and metric
      parts <- strsplit(y, " - ", fixed = TRUE)[[1]]
      indicator_bold <- paste0("<b>", parts[1], "</b>")
      metric <- parts[2]
      
      # Recombine with bold indicator and normal metric
      formatted_label <- paste(indicator_bold, "-", metric)
      
      paste0('<span style="color:', dimension_label_color, ';">', formatted_label, '</span>')
    })
  }
}

# Take formatted figure data, a palette for text colors, the label dimension
# lookup from prepare_figure_data(). Return ggplot graph object
pct_change_graph <- function(figure_data, 
                             dp_text_palette, 
                             label_dimension_lookup) {

  # Renaming input to preserve it for error checking
  data <- figure_data
  
  # Standardize column names
  if ("conf_low" %in% names(data) && "conf_high" %in% names(data)) {
    data <- data %>%
      rename(ci_low = conf_low, ci_high = conf_high)
  }

  data <- data %>%
    mutate(across(c(estimate, ci_low, ci_high), ~ as.numeric(.x))) %>%
    # filter(str_detect(metric, 'easements|wages', negate = TRUE)) %>%
    mutate(
      across(c(estimate, starts_with('ci|conf')), as.numeric),
      outcome = ifelse(is.na(outcome), 'no change', outcome),
      outcome = snakecase::to_title_case(outcome)
    )

  ggplot(data, aes(x = estimate, y = combined_label)) +
    geom_point(aes(color = outcome)) +
    geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
    geom_vline(xintercept = 0) +
    
    # Invisible points to create a second legend (dimension colors)
    geom_point(
      aes(fill = dimension, color = dimension), 
      shape = 22, 
      alpha = 0, 
      show.legend = TRUE
    ) +
    
    # Outcome legend (color of real points and lines)
    scale_color_manual(
      values = c('Better' = 'darkgreen', 'No Change' = 'grey', 'Worse' = 'purple'),
      name = "Outcome"
    ) +
    
    # Dimension legend (fill colors)
    scale_fill_manual(
      values = dp_text_palette,
      name = "Dimension",
      labels = snakecase::to_title_case(names(dp_text_palette))
    ) +
    
    # Add both legends
    guides(
      color = guide_legend(order = 1, nrow = 2),
      fill = guide_legend(
        override.aes = list(shape = 22, size = 4, alpha = 1, color = dp_text_palette),
        order = 2,
        nrow = 2
      )
    ) +
    
    theme_minimal() +
    
    labs(
      x = 'Average Percent Change per Year',
      y = 'Indicator with Metric'
    ) + 
    
    scale_y_discrete(
      labels = make_y_axis_labels(label_dimension_lookup, dp_text_palette)
    ) +
    
    theme(
      legend.position = "top",
      legend.direction = "horizontal",
      legend.title.position = "top",
      text = element_text(family = "Times New Roman"),
      axis.title = element_text(family = "Times New Roman"),
      axis.text = element_text(family = "Times New Roman"),
      legend.title = element_text(family = "Times New Roman", hjust = 0.5),
      legend.text = element_text(family = "Times New Roman"),
      axis.text.y.left = ggtext::element_markdown(),
    )
}
```

This page explores time series data at state and county level, then runs some regressions to see whether metrics are improving or declining over time at the county level. We first run weighted OLS models with robust SEs in @sec-ols. Then in @sec-spdep, we use spatial error models weighted by both geography and weighting variables.

## Time Series

Plot time series of average variable that has at least 2 data points. These are raw, unweighted values.

### County Data

```{r}
#| label: prep_series
#| output: false
# Prepare vector of county variables that are time series
get_str(SMdocs::dp_metrics_county)
vars <- unique(SMdocs::dp_metrics_county$variable_name)
series_vars <- map(vars, ~ {
  count <- SMdocs::dp_metrics_county %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
usethis::use_data(series_vars, overwrite = TRUE)
```

Land value per acre also appears jenky, but it is because Richmond County, NY (Staten Island) has a land value of 2.5 million per acre while most other counties are below 100k. We should probably remove this as an outlier.

```{r}
#| label: county_time_series
#| fig-height: 25
#| fig-width: 12
#| cache: true
#| freeze: true
plots <- map(series_vars, ~ {
  df <- SMdocs::dp_metrics_county %>%
    filter(variable_name == .x) %>%
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```

### State Data

```{r}
#| label: state_time_series
#| fig-height: 12
#| fig-width: 12
vars <- unique(SMdocs::dp_metrics_state$variable_name)
state_series_vars <- map(vars, ~ {
  count <- SMdocs::dp_metrics_state %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
usethis::use_data(state_series_vars, overwrite = TRUE)

plots <- map(state_series_vars, ~ {
  df <- SMdocs::dp_metrics_state %>%
    filter(variable_name == .x) %>%
    mutate(across(c(year, value), as.numeric)) %>% 
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
      # col = '#154734'
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 3,
  nrow = 4
)
```

### Tiny Trend Graphs

Outputs not shown here - these will be added to table in paper.

```{r}
all_series_vars <- c(series_vars, state_series_vars)

# Function to get tiny plots from county and state vars
get_tiny_plots <- function(df, vars) {
  map(vars, ~ {
    df <- df %>%
      filter(variable_name == .x, !is.na(value)) %>% 
      # group_by(variable_name, year) %>% 
      # summarize(annual_mean = mean(value, na.rm = TRUE)) %>% 
      ggplot(aes(x = year, y = value)) +
      geom_line(
        aes(group = fips),
        # col = 'black',
        # size = 0.2,
        alpha = 0.2
      ) +
      geom_smooth(
        method = "loess", se = FALSE, size = 0.75
      ) +
      # geom_hline(
      #   yintercept = 0,
      #   linetype = 'dashed',
      #   color = 'grey'
      # ) +
      coord_cartesian(xlim = c(2000, 2024)) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_discrete(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.margin = margin(rep(0.5, 4))
        
      )
  }) %>% 
    setNames(c(vars))
}

county_plots <- SMdocs::dp_metrics_county %>% 
  # get_normalized_metrics() %>% 
  get_tiny_plots(
    vars = series_vars
  )
county_plots[[1]]
county_plots[[2]]
county_plots[[3]]
county_plots[[10]]
county_plots[[11]]
county_plots[[12]]

# Function to normalize to percent change relative to baseline first year
percent_change <- function(x) {
  x <- as.numeric(x)
  # if (length(x) < 2) stop("Vector must have length >= 2")
  if (sum(!is.na(x)) < 2) {
    warning("Not enough non-NA values to compute percent change; returning NA")
    return(rep(NA_real_, length(x)))
  }
  (x / first(x) - 1) * 100
}


# Normalize by min max or percent change to baseline?
get_normalized_metrics <- function(df) {
  vars <- unique(df$variable_name)
  map(vars, ~ {
    df %>%
      filter(variable_name == .x) %>% 
      mutate(across(c(year, value), as.numeric)) %>% 
      drop_na(value) %>% 
      group_by(variable_name, fips) %>% 
      # mutate(value = min_max(value)) %>%
      mutate(value = percent_change(value)) %>%
      ungroup()
  }) %>% 
    list_rbind()
}

dp_metrics_county %>% 
  arrange(variable_name, fips, year) %>% 
  get_str()
test <- SMdocs::dp_metrics_county %>% 
  get_normalized_metrics()
test %>% 
  arrange(variable_name, fips, year) %>% 
  get_str()

county_plots <- SMdocs::dp_metrics_county %>% 
  # get_normalized_metrics() %>% 
  get_tiny_plots(
    vars = series_vars
  )

county_plots <- SMdocs::dp_metrics_county %>% 
  get_normalized_metrics() %>% 
  get_tiny_plots(
    vars = series_vars
  )

state_plots <- SMdocs::dp_metrics_state %>% 
  get_normalized_metrics() %>% 
  get_tiny_plots(
    vars = state_series_vars
  )

# Save them all to outputs with variable names
iwalk(c(county_plots, state_plots), \(plot, var) {
  ggsave(
    filename = paste0('outputs/trend_plots/fig_trend_', var, '.png'),
    plot = plot,
    width = 1.5,
    height = 1.5,
    units = 'cm',
    dpi = 300,
    bg = NULL
  )
})

# Check milk actual trend
SMdocs::dp_metrics_county %>% 
  group_by(variable_name, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(variable_name == 'retailSalesPropWholesale') %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line()

SMdocs::dp_metrics_state %>% 
  group_by(variable_name, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(variable_name == 'voterTurnout') %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line()

SMdocs::dp_metrics_county %>% 
  group_by(variable_name, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(variable_name == 'totalIndemnities') %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line()

SMdocs::dp_metrics_state %>% 
  group_by(variable_name, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(variable_name == 'yieldMilk') %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line()

SMdocs::dp_metrics_county %>% 
  group_by(variable_name, year) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(variable_name == 'otyAnnualAvgEstabsCountPctChgNAICS11') %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line()

SMdocs::dp_metrics_county %>%
  filter(variable_name == 'otyAnnualAvgEstabsCountPctChgNAICS11') %>% 
  group_by(variable_name, year) %>% 
  mutate(mean_value = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(group = fips), alpha = 0.1) +
  geom_line(aes(x = year, y = mean_value), color = 'red', lwd = 2) +
  geom_smooth() + 
  theme_bw()

SMdocs::dp_metrics_county %>%
  filter(variable_name == 'otyAnnualAvgEstabsCountPctChgNAICS11') %>% 
  group_by(variable_name, year) %>% 
  mutate(
    mean_value = mean(value, na.rm = TRUE),
    pct_change = percent_change(value)
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(group = fips), alpha = 0.1) +
  geom_line(aes(x = year, y = mean_value), color = 'red', lwd = 2) +
  geom_line(aes(x = year, y = pct_change), color = 'darkgreen', lwd = 2) +
  geom_smooth() +
  theme_bw()

###
df <- SMdocs::dp_metrics_county %>%
  filter(variable_name == 'otyAnnualAvgEstabsCountPctChgNAICS11') %>% 
  group_by(variable_name, fips) %>% 
  arrange(year) %>%
  mutate(pct_change = percent_change(value)) %>% 
  ungroup()

df_avg <- df %>%
  group_by(variable_name, year) %>%
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    mean_pct_change = mean(pct_change, na.rm = TRUE),
    .groups = "drop"
  )

ggplot() +
  geom_line(data = df, aes(x = year, y = value, group = fips), alpha = 0.1) +
  geom_line(data = df_avg, aes(x = year, y = mean_value), color = 'red', lwd = 1.2) +
  geom_line(data = df_avg, aes(x = year, y = mean_pct_change), color = 'darkgreen', lwd = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", colo = "gray") +
  theme_bw()
```

## Linear Models {#sec-ols}

First trying linear models with time as independent variable. Min max normalizing dataset with stored function before we do anything else.

```{r}
#| label: normalize
#| output: false
get_str(SMdocs::dp_metrics_county)

vars <- unique(SMdocs::dp_metrics_county$variable_name)
dp_metrics_county_normed <- map(vars, ~ {
  SMdocs::dp_metrics_county %>%
    filter(variable_name == .x) %>% 
    mutate(across(c(year, value), as.numeric)) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
get_str(dp_metrics_county_normed)  
usethis::use_data(dp_metrics_county_normed, overwrite = TRUE)
```

### Regressions

First create a weights crosswalk that links `variable_name`s to weighting variables.

```{r}
#| label: weights_crosswalk
#| output: false
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Population (5 year, smooth)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres operated, smooth',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5Year',
    'population5YearSmooth',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'acresOperatedSmooth',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)
weights_crosswalk

# Link weight vars to the appropriate variables
weights_crosswalk <- SMdocs::dp_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% SMdocs::dp_metrics_county_normed$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric) %>% 
  na.omit()
get_str(weights_crosswalk)
weights_crosswalk
```

Run regressions for all time series variables where time points \> 1.

```{r}
#| label: regressions
#| output: false
#| eval: false
# Get a crosswalk to add state as a fixed effect
state_link <- SMdata::fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

res <- map(SMdocs::series_vars, ~ {
  tryCatch(
    {
      df <- SMdocs::dp_metrics_county_normed %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year),
          state_fips = str_sub(fips, end = 2)
        ) %>% 
        left_join(state_link, by = join_by(state_fips == fips))
      
      # Add weighting variable
      weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == .x]
      weight_df <- SMdocs::dp_weights
      if (!is.na(weight_var) && length(weight_var) > 0) {
        weight_df <- weight_df %>% 
          filter(variable_name == weight_var)
      }
      weight_df <- weight_df %>% 
        mutate(
          value = as.numeric(value), 
          year = as.numeric(year)
        ) %>% 
        rename(
          weight_value = value, 
          weight_name = variable_name,
          weight_year = year
        )
      # get_str(weight_df)
      
      # Depending on whether weight has a year, do appropriate join
      # Note that we might be losing some years here if weighting var doesn't
      # cover the variable enough
      if (all(is.na(weight_df$weight_year))) {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips))
      } else {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips, year == weight_year))
      }
      
      # Reduce to county fips only. Remove dupes 
      df <- df %>%
        filter(str_length(fips) == 5) %>% 
        distinct()
      
      # Weighted OLS 
      model <- lm(
        normed_value ~ year + state_code, 
        data = df,
        weights = weight_value
      )
      return(model)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(SMdocs::series_vars)) %>%
  discard(\(x) is.null(x))
get_str(res, 1)
get_str(res)

# Robust SEs
lmtest::coeftest(res[[1]])

# Save immediate output
saveRDS(res, 'data/objects/weighted_ols_outs.rds')
```

### Residuals

Check residuals of outputs:

```{r}
#| label: lm_residuals
#| freeze: true
#| warnings: false
#| outputs: false
#| eval: false
# res <- readRDS('data/objects/weighted_ols_outs.rds')
# res_plots <- map(res, ~ {
#   try(performance::check_model(.x))
# }) %>% 
#   purrr::keep(~ length(.x) > 1)
# saveRDS(res_plots, 'data/objects/weighted_ols_residuals.rds')
```

```{r}
#| fig-width: 10
#| fig-height: 10
#| eval: false
#| output: false
#| echo: false
# res_plots <- readRDS('data/objects/weighted_ols_residuals.rds')
# res_plots
```

(Residual plots are giving us trouble currently)

Now that we have regression outputs, let's put them in a table to see which ones change significantly over time.

```{r}
#| label: make_reg_table
#| output: false
# Put together one table of results from all models, add stars
res <- readRDS('data/objects/weighted_ols_outs.rds')
get_str(res)

# Check for NA outputs
res[[1]]$coefficients[['year']]
coefs <- map(res, ~ .x$coefficients[['year']])
names(coefs[is.na(coefs)])

# Mobility disability is NA because there is only 1 year after we join with 
# weighting variable. So we need to just throw it out if we can't get smoothed
# population for 2024.
res <- res %>% 
  discard(~ is.na(.x$coefficients[[2]]))

# Make a table of regression outputs, including robust standard errors
res_df <- imap(res, ~ {
  vcov = vcovHC(.x, type = 'HC0')
  df <- coeftest(.x, vcov = vcov) %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
  conf_ints <- coefci(.x)['year', ]
  df$ci_low <- conf_ints[1]
  df$ci_high <- conf_ints[2]
  df
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many time series metrics total
(n_metrics <- nrow(res_df))

# How many change over time significantly
(
  n_change <- res_df %>% 
    filter(sig == '*') %>% 
    nrow()
)

# Percentage significant
(perc_sig <- mean(res_df$sig == '*', na.rm = TRUE) * 100)
```

Note that we only have 2 years for the mobility disability metric - 2023 and 2024. But we don't have a smoothed 5-year population for 2024, only 2023. So we can only use a single year, which means there is no variation in year. So we are dropping this from the analysis.

Out of `{r} n_metrics` metrics with \> 1 time point, we have `{r} n_change` metrics that change significantly over time (`{r} round(perc_sig, 2)`%).

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch.

```{r}
#| label: make_sense_of_trends
#| output: false
# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
raw_outputs <- res_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(SMdocs::dp_meta)
directions <- SMdocs::dp_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics
model_outputs <- raw_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# See what we lost
(lost <- setdiff(raw_outputs$term, model_outputs$term))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).

Check out table with outputs for each regression. Filter by significance and outcome. Trend is just whether it is going up or down, outcome is whether that is good or bad.

### Table

```{r}
#| label: reg_table
# Add framework info
framed_outputs <- SMdocs::dp_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(model_outputs, by = join_by(variable_name == term))

# Create interactive table for Quarto
get_reactable(
  framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```

### Figure

Figure below shows 95% confidence intervals for the coefficients of each metric, including state as a covariate.

```{r}
#| label: regression_figure
# Combined label for indicator and metric
# output_data <- framed_outputs %>%
#   mutate(
#     indicator = snakecase::to_title_case(indicator),
#     metric = snakecase::to_title_case(metric),
#     combined_label = paste(indicator, "-", metric)  # Combine indicator and metric
#   )
  
# Color code by dimension
# dimension_colors <- c(
#   "economics"  = "#BF7417FF",
#   "environment"  = "#A8BE74FF",
#   "health" = "#9CB6A9FF",
#   "production" = "#635C72FF",
#   "social" = "#784116FF"
# )

# # Order dimensions alphabetically
# ordered_dims <- sort(unique(output_data$dimension))
# 
# # Group by dimension, then alphabetically, and filter out "poor mental health"
# output_data <- output_data %>%
#   mutate(
#     dimension = factor(dimension, levels = ordered_dims)
#   ) %>%
#   arrange(dimension, estimate) %>%
#   mutate(
#     combined_label = factor(combined_label, levels = rev(unique(combined_label))) 
#   ) %>%
#   filter(metric != "Poor Mental Health Days") %>%
#   ungroup()
# 
# # Getting colors correctly mapped to combined labels
# label_dimension_lookup <- output_data %>%
#   distinct(combined_label, dimension)
# 
# # Percent change graph
# lm_figure <- output_data %>% 
#   mutate(
#     across(c(estimate, starts_with('ci')), as.numeric),
#     outcome = ifelse(is.na(outcome), 'no change', outcome) %>% 
#       snakecase::to_title_case()
#   ) %>% 
#   ggplot(aes(x = estimate, y = combined_label)) +
#   geom_point(aes(color = outcome)) +
#   geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
#   geom_vline(xintercept = 0) +
#   
#   # Invisible points to create a second legend (dimension colors)
#   geom_point(
#     aes(fill = dimension, color = dimension), 
#     shape = 22, 
#     alpha = 0, 
#     show.legend = TRUE
#   ) +
#   
#   # Outcome legend (color of real points and lines)
#   scale_color_manual(
#     values = c('Better' = 'darkgreen', 'No Change' = 'grey', 'Worse' = 'purple'),
#     name = "Outcome"
#   ) +
#   
#   # Dimension legend (fill colors)
#   scale_fill_manual(
#     values = dp_text_palette,
#     name = "Dimension",
#     labels = snakecase::to_title_case(names(dp_text_palette))
#   ) +
#   
#   # Add both legends
#   guides(
#     color = guide_legend(order = 1, nrow = 2),
#     fill = guide_legend(
#       override.aes = list(shape = 22, size = 4, alpha = 1, color = dp_text_palette),
#       order = 2,
#       nrow = 2
#     )
#   ) +
#   
#   theme_minimal() +
#   
#   labs(
#     x = 'Average Percent Change per Year',
#     y = 'Indicator with Metric'
#   ) + 
#   
#   scale_y_discrete(
#     labels = function(x) {
#       sapply(x, function(y) {
#         dim_val <- label_dimension_lookup$dimension[label_dimension_lookup$combined_label == y]
#         dimension_label_color <- dp_text_palette[as.character(dim_val)]
#         
#         # Split the combined label into indicator and metric
#         parts <- strsplit(y, " - ", fixed = TRUE)[[1]]
#         indicator_bold <- paste0("<b>", parts[1], "</b>")
#         metric <- parts[2]
#         
#         # Recombine with bold indicator and normal metric
#         formatted_label <- paste(indicator_bold, "-", metric)
#         
#         paste0('<span style="color:', dimension_label_color, ';">', formatted_label, '</span>')
#       })
#     }
#   ) +
#   theme(
#     legend.position = "top",
#     legend.direction = "horizontal",
#     legend.title.position = "top",
#     text = element_text(family = "Times New Roman"),
#     axis.title = element_text(family = "Times New Roman"),
#     axis.text = element_text(family = "Times New Roman"),
#     legend.title = element_text(family = "Times New Roman", hjust = 0.5),
#     legend.text = element_text(family = "Times New Roman"),
#     axis.text.y.left = ggtext::element_markdown(),
#   )

processed_lm_data <- prepare_figure_data(framed_outputs)

lm_figure <- pct_change_graph(
  figure_data = processed_lm_data$figure_data,
  dp_text_palette = dp_text_palette,
  label_dimension_lookup = processed_lm_data$label_dimension_lookup
)

ggsave(
  'outputs/fig_lm.png',
  plot = lm_figure,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)

lm_figure
```

## Spatial Regressions {#sec-spdep}

### One Model

`spatialreg` for spatially weighted regressions, but not panels. First trying a single example with gini. Note that because Census data spans the change in administrative boundaries in CT, we can't use CT at all here.

```{r}
#| label: spatialreg_test
#| output: false
get_str(SMdocs::dp_metrics_county)

dat <- SMdocs::dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  )
get_str(dat)

# Remove CT. Drop na ginis, but, get complete cases for relevant fips and years
dat <- dat %>% 
  select(fips, year, gini) %>% 
  mutate(
    across(!fips, as.numeric),
    gini_normed = min_max(gini)
  ) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>%
  drop_na(gini) %>% 
  complete(fips, year)
get_str(dat)

# List weights
lw <- neast_counties_2024 %>% 
  right_join(dat) %>% 
  select(fips, year, gini) %>% 
  poly2nb() %>% 
  nb2listw(zero.policy = TRUE)

# Regular weights
get_str(weights_crosswalk)
weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == 'gini']
weight_df <- SMdocs::dp_weights

weight_df <- weight_df %>% 
  mutate(
    value = as.numeric(value), 
    year = as.numeric(year)
  ) %>% 
  rename(
    weight_value = value, 
    weight_name = variable_name,
    weight_year = year
  ) %>% 
  filter(weight_name == weight_var, str_length(fips) == 5)
get_str(weight_df)

get_str(dat)
dat <- left_join(dat, weight_df, by = join_by(fips, year == weight_year)) %>% 
  distinct()
get_str(dat)

# Run regression
sar <- errorsarlm(
  gini_normed ~ year,
  data = dat,
  listw = lw,
  weight = weight_value
)
```

```{r}
summary(sar)
```

### All Models

Use `errorsarlm` on all metrics.

```{r}
#| label: spatial_reg_all
#| eval: false
#| output: false
get_str(SMdocs::dp_metrics_county)
state_link <- SMdata::fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

dat <- SMdocs::dp_metrics_county %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  ) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>% 
  mutate(link = str_sub(fips, end = 2)) %>% 
  left_join(state_link, by = join_by(link == fips)) %>% 
  select(-link) %>% 
  mutate(across(!c(fips, state_code), as.numeric))
get_str(dat)

# Get vector of vars
(vars <- names(dat)[!names(dat) %in% c('fips', 'year', 'state_code')])

tic()
get_time()
plan(multisession, workers = availableCores(omit = 2))
# out <- map(vars[1], \(var) {
out <- future_map(vars, \(var) {
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var), state_code) %>% 
      drop_na(!!sym(var)) %>% 
      mutate(normed = min_max(!!sym(var)))
 
    # Regular weights
    weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == var]
    if (!is.na(weight_var) && length(weight_var) > 0) {
      weight_df <- SMdocs::dp_weights %>% 
        mutate(
          value = as.numeric(value), 
          year = as.numeric(year)
        ) %>% 
        rename(
          weight_value = value, 
          weight_name = variable_name,
          weight_year = year
        ) %>% 
        filter(weight_name == weight_var, str_length(fips) == 5)
      
      # Join weights back to df
      # df <- df %>% 
      #   inner_join(weight_df, by = join_by(fips, year == weight_year))
      if (all(is.na(weight_df$weight_year))) {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips))
      } else {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips, year == weight_year))
      }
      
      cat('\nJoined to weights:', weight_var)
    } else {
      cat('\nNo weighting variable')
    }
    
    # List weights
    lw <- SMdata::neast_counties_2024 %>% 
      right_join(df) %>% 
      select(fips, year, normed) %>% 
      poly2nb() %>% 
      nb2listw(zero.policy = TRUE)   
    cat('\nCreated lw')
    
    # Assign weights arg if variable is weighted. Otherwise NULL
    if (!is.na(weight_var) && length(weight_var) > 0) {
      weights_arg <- df[['weight_value']]
    } else {
      weights_arg <- NULL
    }
    
    # Run error sar model with appropriate (or no) weighting
    model <- errorsarlm(
      normed ~ as.numeric(year) + state_code,
      data = df,
      listw = lw,
      weights = weights_arg
    ) 
    cat('\nRan model')
   
    return(model)
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
}, .progress = FALSE) %>% 
  setNames(c(vars))
plan(sequential)
toc()
# 6 to 12 minutes or so

# out
# out[[1]]

get_str(out)
get_str(out[[1]])
map(out, class)
# get_str(out[[1]])
# map(out, ~ .x['fips'])
# map(out, ~ summary(.x))

saveRDS(out, 'data/objects/errorsarlm_outs.rds')
```

#### Reactable Table

```{r}
#| output: false
sar_outs <- readRDS('data/objects/errorsarlm_outs.rds')
start <- length(sar_outs)
sar_outs <- sar_outs[names(sar_outs) != 'state_code'] %>% 
  keep(~ class(.x) == 'Sarlm')
end <- length(sar_outs)
(lost <- start - end)
# 42 down to 38 here

sar_df <- imap(sar_outs, ~ {
  tryCatch(
    {
      df <- tidy(.x)
      est <- df[2, 'estimate'] %>% as.numeric
      se <- df[2, 'std.error'] %>% as.numeric
      alpha <- 0.05
      z <- qnorm(1 - alpha/2)
      confints <- c(est - z * se, est + z * se)
      
      # confints <- confint(.x)[3, ]
      out <- df %>% 
        filter(str_detect(term, 'year')) %>% 
        mutate(
          variable_name = .y,
          conf_low = confints[1],
          conf_high = confints[2]
        ) %>% 
        select(-term) %>% 
        mutate(sig = ifelse(p.value < 0.05, '*', ' ') %>% as.character())  
      return(out)
    },
      error = function(e) {
        message('Error')
        print(e)
    }
  )
}) %>% 
  keep(is.data.frame) %>% 
  bind_rows()
get_str(sar_df)
# sar_df

# 38 down to 35.
(lost <- setdiff(names(sar_outs), sar_df$variable_name))
check <- sar_outs[names(sar_outs) %in% lost]
# Not defined because of singularities. Single time points. 
# At least while including the weighting variables.

sar_raw_outputs <- sar_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))
get_str(sar_raw_outputs)
# sar_raw_outputs

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(SMdocs::dp_meta)
directions <- SMdocs::dp_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics

sar_model_outputs <- sar_raw_outputs %>% 
  filter(variable_name %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(variable_name %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    outcome = ifelse(sig != '*', NA, outcome),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
sar_model_outputs

(n_metrics <- dim(sar_model_outputs)[1])

# See what we lost
(lost <- setdiff(sar_raw_outputs$variable_name, sar_model_outputs$variable_name))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))

# proportion getting better or worse
(tab <- get_table(sar_model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

We have `{r} n_metrics` metrics modeled. `{r} percs[['better']]`% are getting better, `{r} percs[['worse']]`% are getting worse, and `{r} percs[3]`% are stable. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).

```{r}
#| label: errorsar_reactable
sar_framed_outputs <- SMdocs::dp_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(sar_model_outputs, by = join_by(variable_name == variable_name))

## Create interactive table for Quarto
get_reactable(
  sar_framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```

#### LaTeX Table

Saving table to $\LaTeX$ for paper:

```{r}
#| label: latex_table
#| output: false
get_str(sar_framed_outputs)

# Reduce to specific columns
latex_df <- sar_framed_outputs %>% 
  select(metric, estimate:p.value, sig, outcome) %>% 
  mutate(
    p.value = ifelse(sig == '*', paste0(p.value, '*'), p.value),
    outcome = case_when(
      is.na(outcome) ~ '\\textemdash',
      .default = str_to_title(outcome)
    )
  ) %>% 
  select(-sig) %>% 
  setNames(c(
    'Metric',
    '$\\beta$',
    'SE',
    '$z$-value',
    'p-value',
    'Outcome'
  ))
get_str(latex_df)

# Save to latex
latex_df %>% 
  kbl(
    format = 'latex',
    caption = 'Coefficients for spatial error models of year on metric',
    label = 'tab_errorsarlm',
    booktabs = TRUE,
    escape = FALSE
  ) %>% 
  kable_styling(
    font_size = 9
  ) %>% 
  footnote(
    general_title = "",
    general = 
      '\\\\textit{Note: }Table shows a series of spatial error models with year
      as and state as predictor and metric as the dependent variable, weighted
      by variables shown in \\\\ref{tab:tab_metrics_body}. Significant p-values
      indicated change over time, and the direction of the outcome was determined
      based on desirable directions of change for each metric.',
    threeparttable = TRUE,
    escape = FALSE
  ) %>% 
  save_kable(
    file = 'outputs/tab_errorsarlm.tex'
  )
```

#### Figure

```{r}
#| label: errorsar_figure
# Using functions at top of script to process outputs and make graph
errorsar_figure_data <- prepare_figure_data(model_output_df = sar_framed_outputs)
errorsar_plot <- pct_change_graph(
  figure_data = errorsar_figure_data$figure_data,
  dp_text_palette = dp_text_palette,
  label_dimension_lookup = errorsar_figure_data$label_dimension_lookup
)

ggsave(
  'outputs/fig_errorsarlm.png',
  plot = errorsar_figure,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)

errorsar_figure
```

