---
title: "Indicator Distributions"
format:
  html:
    fig-dpi: 200
warnings: false
---

```{r}
#| label: prep
#| echo: false
pacman::p_load(
  dplyr,
  conflicted
)
pacman::p_load_gh('ChrisDonovan307/projecter')
conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::summarize(),
  .quiet = TRUE
)
```

Now that our metrics have been aggregated into indicators, we can observe the univariate distributions of the indicators themselves. We will do this for all six sets of transformations.

# Summary

Below are counts of skewed indicator distributions for each transformation. 

::: {style="display: flex; justify-content: center;"}
```{r}
#| label: summary
#| fig.align: center
pacman::p_load(
  purrr,
  dplyr,
  tibble
)

# Load reactable custom function
source('dev/get_reactable.R')

# Load scores data for all transformations
scores <- readRDS('data/state_score_iterations.rds')
# get_str(scores)

# Rearrange data to make a single DF as table
out <- map_vec(scores, ~ {
  .x[['indicator_scores']] %>% 
    filter(!state %in% c('US_mean', 'US_median', 'NewEng')) %>% 
    select(-state) %>% 
    psych::describe() %>% 
    select(skew) %>% 
    filter(abs(skew) > 2) %>% 
    nrow()
}) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  setNames(c('Transformation', 'Indicators with Skew > 2')) %>% 
  arrange(Transformation)
# out

get_reactable(
  out, 
  fullWidth = FALSE,
  searchable = FALSE,
  defaultColDef = colDef(
    minWidth = 200
  )
)
```
:::

It looks like the Box Cox procedure transformed distributions at the metric level enough that there are no skewed distributions at the indicator level. Both the min-max and z-score transformations left us with only three skewed indicators: access to culturally appropriate food, total quantity exported, and total quantity imported. The latter two are not surprising, as big agricultural states likely have a disproportionate impact. The Z-score with geometric aggregation is a slightly nonsensical combination of methods that we will ignore for now.

This is providing some evidence that Box-Cox transformations could be warranted on at least these three problematic indicators (or the metrics therein). Below are the indicator distributions for each transformation. Inidicators with skew > 2 are shown in red.

# Min Max Arithmetic

```{r}
#| label: minmax_arithmetic
#| fig-cap: Distributions of indicators at the state level for the min-max arithmetic transformations.
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
pacman::p_load(
  ggplot2,
  purrr,
  ggpubr
)

# Load custom function
source('dev/get_indicator_distributions.R')

# Plot them
get_indicator_distributions(scores, 'minmax_arithmetic')
```

# Min Max Geometric

```{r}
#| label: minmax_geo
#| fig-cap: Distributions of indicators at the state level for the min-max geometric transformations
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'minmax_geometric')
```

# Z-Score Arithmetic

```{r}
#| label: zscore_arithmetic
#| fig-cap: Distributions of indicators at the state level for the Z-score arithmetic transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'zscore_arithmetic')
```

# Z-Score Geometric

```{r}
#| label: zscore_geometric
#| fig-cap: Distributions of indicators at the state level for the Z-score geometric transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'zscore_geometric')
```

# Box Cox Arithmetic

```{r}
#| label: boxcox_arithmetic
#| fig-cap: Distributions of indicators at the state level for the Box Cox arithmetic transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'boxcox_arithmetic')
```

# Box Cox Geometric

```{r}
#| label: boxcox_geometric
#| fig-cap: Distributions of indicators at the state level for the Box Cox geometric transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'boxcox_geometric')
```

# Rank Arithmetic

```{r}
#| label: rank_arithmetic
#| fig-cap: Distributions of indicators at the state level for the rank arithmetic transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'rank_arithmetic')
```

# Rank Geometric

```{r}
#| label: rank_geometric
#| fig-cap: Distributions of indicators at the state level for the rank geometric transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'rank_geometric')
```


# Winsor Arithmetic

```{r}
#| label: winsor_arithmetic
#| fig-cap: Distributions of indicators at the state level for the winsor arithmetic transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'winsor_arithmetic')
```

# Winsor Geometric

```{r}
#| label: winsor_geometric
#| fig-cap: Distributions of indicators at the state level for the winsor geometric transformations 
#| fig-height: 15
#| fig-width: 10
#| fig-align: center
get_indicator_distributions(scores, 'winsor_geometric')
```
