---
title: "Metrics Database"
---

```{r}
#| label: prep
#| warnings: false
#| echo: false
#| cache: false
devtools::load_all()
get_setup()

pacman::p_load(
  dplyr,
  stringr,
  readr
)

# Combine fips key and state key to identify all fips codes
slim_state_key <- SMdata::state_key %>% 
  select(fips = state_code, name = state_name)
slim_fips_key <- SMdata::fips_key %>% 
  select(fips, name = county_name) %>% 
  filter(str_length(fips) == 5)
all_fips_key <- bind_rows(slim_state_key, slim_fips_key)
write_csv(all_fips_key, 'data/all_fips_key.csv')
```

::: {.callout-note collapse='false' title='Caution'}
The SM database is still a work in progress. It has not been updated to reflect new metrics gathered in the Northeast regional data survey, nor has it been optimized to work particularly well. In the future, we will be fixing these issues and adding public API endpoints to make downloading data a more pleasant experience. 
:::

Use the interactive app below to filter for metrics by dimension, index, and indicator. Dropdown lists will adjust automatically. Note that large queries will take a few seconds to retrieve, mostly because of adding all the associated metadata. Will try to sort that out eventually. Use the query button to send a query to the database and the download button to download a .csv of the current, filtered table.

```{r}
#| label: metrics_bulk_download
#| include: false
#| eval: false
pacman::p_load(
  dplyr,
  reactable,
  stringr,
  htmltools,
  readr
)

metrics_file_name = paste0(Sys.Date(), "_bulk_metrics.csv")
fips_file_name = paste0(Sys.Date(), "_fips_key.csv")

tagList(
  tags$div(
    style = "display: flex; gap: 100px; margin-bottom: 20px; justify-content: center;",
    tags$a(
      class = "btn btn-primary",
      style = "display: flex; width: 200px; justify-content: center; align-items: center;",
      href = '../data/bulk_metrics.csv',
      download = metrics_file_name,
      tagList(fontawesome::fa("download"), "Download Bulk .CSV")
    ),
   tags$a(
      class = "btn btn-primary",
      style = "display: flex; width: 200px; justify-content: center; align-items: center;",
      href = '../data/all_fips_key.csv',
      download = fips_file_name,
      tagList(fontawesome::fa("download"), "Download FIPS Key")
    )
  )
)

```

<iframe width="1200" height="1000" src="https://fsrc.w3.uvm.edu/SMdb/" title="Secondary Metrics Database"></iframe>
