---
title: "Metrics Table"
filters:
  - shinylive
execute:
  eval: false
---

```{r}
#| label: setup
#| warnings: false
#| echo: false
pacman::p_load(
  conflicted
)
conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  .quiet = TRUE
)
```

This page contains a metadata table for exploring the sources of secondary data used throughout this project.

Using the table: 

* Click column headers to sort
* Global search in the top right, or column search in each header
* Change page length and page through results at the bottom
* Use the download button to download a .csv file of the filtered table
* Click the arrow on the left of each row for details, including a URL to the data source.

```{r}
#| label: metrics_table
#| warning: false

pacman::p_load(
  dplyr,
  reactable,
  stringr,
  htmltools
)

# Load full metadata table
metadata_all <- readRDS('data/sm_data.rds')[['metadata']]

# Pick out variables to display
metadata <- metadata_all %>% 
  select(
    metric,
    'Variable Name' = variable_name,
    definition,
    dimension,
    index,
    indicator,
    units,
    'Year' = latest_year, # Renaming latest year as year, not including og year
    source,
    scope,
    resolution,
    url
) %>% 
  setNames(c(str_to_title(names(.))))

###
htmltools::browsable(
  tagList(
    
    tags$div(
      style = "display: flex; gap: 16px; margin-bottom: 20px; justify-content: center;",
      
      tags$button(
        class = "btn btn-primary",
        style = "display: flex; align-items: center; gap: 8px; padding: 8px 12px;",
        tagList(fontawesome::fa("download"), "Show/hide more columns"),
        onclick = "Reactable.setHiddenColumns('metadata_table', prevColumns => {
          return prevColumns.length === 0 ? ['Definition', 'Scope', 'Resolution', 'Url'] : []
        })"
      ),
      
      tags$button(
        class = "btn btn-primary",
        style = "display: flex; align-items: center; gap: 8px; padding: 8px 12px;",
        tagList(fontawesome::fa("download"), "Download as CSV"),
        onclick = "Reactable.downloadDataCSV('metadata_table', 'sustainability_metadata.csv')"
      )
    ),
    
    reactable(
      metadata,
      sortable = TRUE,
      resizable = TRUE,
      filterable = TRUE,
      searchable = TRUE,
      pagination = TRUE,
      bordered = TRUE,
      wrap = TRUE,
      rownames = FALSE,
      onClick = 'select',
      striped = TRUE,
      pageSizeOptions = c(5, 10, 25, 50, 100),
      defaultPageSize = 5,
      showPageSizeOptions = TRUE,
      highlight = TRUE,
      style = list(fontSize = "14px"),
      compact = TRUE,
      fullWidth = TRUE,
      columns = list(
        Metric = colDef(
          minWidth = 200,
          sticky = 'left'
        ),
        'Variable Name' = colDef(
          minWidth = 150
        ),
        Definition = colDef(
          minWidth = 250
        ),
        'Latest Year' = colDef(minWidth = 75),
        Source = colDef(minWidth = 250),
        Scope = colDef(show = FALSE),
        Resolution = colDef(show = FALSE),
        Url = colDef(
          minWidth = 300,
          show = FALSE
        )
      ),
      defaultColDef = colDef(minWidth = 100),
      elementId = "metadata_table",
      details = function(index) {
        div(
          style = "padding: 15px; border: 1px solid #ddd; margin: 10px 0;
             background-color: #E0EEEE; border-radius: 10px; border-color: black;
             box-shadow: 2px 2px 10px rgba(0, 0, 0, 0.1);",
          
          tags$h4(
            strong("Details"), 
          ),
          tags$p(
            strong('Metric Name: '), 
            as.character(metadata_all[index, 'metric']),
          ),
          tags$p(
            strong('Variable Name: '), 
            as.character(metadata_all[index, 'variable_name']),
          ),
          tags$p(
            strong('Definition: '), 
            as.character(metadata_all[index, 'definition']),
          ),
          tags$p(
            strong('Source: '), 
            as.character(metadata_all[index, 'source'])
          ),
          tags$p(
            strong('Latest Year: '), 
            as.character(metadata_all[index, 'latest_year'])
          ),
          tags$p(
            strong('All Years (cleaned, wrangled, and included): '), 
            as.character(metadata_all[index, 'year'])
          ),
          tags$p(
            strong('Updates: '), 
            str_to_title(as.character(metadata_all[index, 'updates']))
          ),
          tags$p(
            strong('URL: '), 
            tags$a(
              href = as.character(metadata_all[index, 'url']),
              target = '_blank',
              as.character(metadata_all[index, 'url'])
            )
          )
        )
      }
    )
  )
)

```

